---
author: "E.E. Vorsteveld"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Packages and functions
```{r}
library(tidyverse)
library(here)
library(venn)
library(rlist)
library(DESeq2)
library(DEFormats)
library(reshape2)
library(corrplot)
library(RColorBrewer)
library(factoextra)
library(limma)
library(msigdbr)
library(fgsea)
library(org.Hs.eg.db)
library(glue)
library(ggrepel)
library(ggpubr)
library(correlationAnalyzeR)
library(matrixcalc)
library(scales)
library(gtools)
library(ggrepel)
library(ggh4x)
library(UpSetR)
library(ggtext)

source(here::here("functions/fgseaFromdds.R"))
source(here::here("functions/fisherGeneList.R"))
source(here::here("functions/jaccard.R"))
source(here::here("functions/mRNA_function.R"))
source(here::here("functions/getGenes.R"))
source(here::here("functions/fixTimeOrder.R"))
source(here::here("functions/fixClassOrder.R"))
source(here::here("functions/fixConditionOrder.R"))
source(here::here("functions/getDataForGSEAPlot.R"))
source(here::here("functions/GSEASimple.R"))
source(here::here("functions/addTranscript.R"))
```

# Setup
Metadata
```{r}
seq_data <- here::here("data/seq_data")

#Annotation file available from ensembl: https://www.ensembl.org/Homo_sapiens/Info/Index
ref_annot <- read.table(here("data/Homo_sapiens.GRCh38.95_nodupes_genid.txt"), header = TRUE, sep = "\t")

metadat <- read.csv(here("data/Phenodata.txt"), header = TRUE, sep = "\t")
row.names(metadat) <- metadat$id
```

Generating a DESeq dataset
```{r}
sampleTable_4hr <- data.frame(sampleName = (metadat[grep("*_4hr", metadat$id), ]$id), 
                              fileName = (gsub("$", ".counts.txt", 
                                               (metadat[grep("*_4hr", metadat$id), ]$id))), 
                              stimulus = (metadat[grep("*_4hr", metadat$id), ]$stimulus), 
                              donor = (metadat[grep("*_4hr", metadat$id), ]$donor), 
                              time_point = (metadat[grep("*_4hr", metadat$id), ]$time))

sampleTable_24hr <- data.frame(sampleName = (metadat[grep("*_24hr", metadat$id), ]$id), 
                               fileName = (gsub("$", ".counts.txt", 
                                                (metadat[grep("*_24hr", metadat$id), ]$id))), 
                               stimulus = (metadat[grep("*_24hr", metadat$id), ]$stimulus), 
                               donor = (metadat[grep("*_24hr", metadat$id), ]$donor), 
                               time_point = (metadat[grep("*_24hr", metadat$id), ]$time))

sampleTable_list <- list()
sampleTable_list <- rlist::list.append(sampleTable_list, "sampleTable_4hr")
sampleTable_list <- rlist::list.append(sampleTable_list, "sampleTable_24hr")

opt.time <- c("4hr", "24hr")
#opt.donors <- c("A", "B", "C", "D", "E")

ddsHTSeq_list <- list()
ddsHTSeq_list <- rlist::list.append(ddsHTSeq_list, "ddsHTSeq_4hrs")
ddsHTSeq_list <- rlist::list.append(ddsHTSeq_list, "ddsHTSeq_24hrs")

makeDESeqDataSet <- function(t){
  DESeqDataSetFromHTSeqCount(sampleTable = t, directory = seq_data, design = ~ donor + stimulus)
}

for (i in 1:length(sampleTable_list)){
  assign(as.character(ddsHTSeq_list[[i]][1]),
         makeDESeqDataSet(
           get(as.character(sampleTable_list[[i]][1]), envir = globalenv())
           )
         )
}
```

Filter based on rowsums
```{r}
keepFromddsHTSeq <- function(t){
  keep <- rowSums(counts(t)) >= 10
  t[keep, ]
}

for (i in 1:length(ddsHTSeq_list)){
  assign(as.character(ddsHTSeq_list[[i]][1]),
         keepFromddsHTSeq(
           get(as.character(ddsHTSeq_list[[i]][1]), envir = globalenv()))
         )
}
```

Remove mtrRNA, mttRNA, snRNA, snoRNA, miscRNA and lincRNA genes
```{r}
#Apply the mRNA filtering function to the list of ddsHTSeq objects
for (i in 1:length(ddsHTSeq_list)){
  assign(as.character(ddsHTSeq_list[[i]][1]),
         mRNA_function(get(
           as.character(ddsHTSeq_list[[i]][1]), envir = globalenv())
                       )
         )
} 
```

Differential expression analysis
```{r}
source(here::here("functions/diffExpression.R"))

dds_list <- list()
dds_list <- list.append(dds_list, "dds_4hr")
dds_list <- list.append(dds_list, "dds_24hr")

#Apply the diffExpression function to the list of ddsHTSeq objects
for (i in 1:length(ddsHTSeq_list)){
  assign(as.character(dds_list[[i]][1]),
         diffExpression(get(as.character(ddsHTSeq_list[[i]][1]), envir = globalenv()))
         )
}
```

log-fold shrinkage of the result from DESeq
```{r}
source(here("functions/logFoldShrinkage.R"))
library(apeglm)

opt.cond.long <- c("group_LPS_vs_RPMI", 
                   "group_S_Aureus_vs_RPMI", 
                   "group_PolyIC_vs_RPMI", 
                   "group_C_Albicans_vs_RPMI")

saveRDS(opt.cond.long, here("Donor/opt_cond_long.rda"))
opt.cond.long <- readRDS(here("Donor/opt_cond_long.rda"))

resLFC4hrLPS <-         logFoldShrinkage(dds_4hr, opt.cond.long[1])
resLFC4hrSAur <-        logFoldShrinkage(dds_4hr, opt.cond.long[2])
resLFC4hrPolIC <-       logFoldShrinkage(dds_4hr, opt.cond.long[3])
resLFC4hrCAlb <-        logFoldShrinkage(dds_4hr, opt.cond.long[4])

resLFC24hrLPS <-        logFoldShrinkage(dds_24hr, opt.cond.long[1])
resLFC24hrSAur <-       logFoldShrinkage(dds_24hr, opt.cond.long[2])
resLFC24hrPolIC <-      logFoldShrinkage(dds_24hr, opt.cond.long[3])
resLFC24hrCAlb <-       logFoldShrinkage(dds_24hr, opt.cond.long[4])
```

Generating a masterfile
```{r}
opt.time <- c("4hr", "24hr")
resLFC_list <- list()

for (time in opt.time){
  for (cond in opt.cond){ 
  name <- as.character(paste0("resLFC", time, cond))

  resLFC_list <- list.append(name, resLFC_list)

  t <- get(name)
  assign(name,
         as.data.frame(t) %>% 
           mutate(time = time) %>% 
           mutate(condition = cond) %>% 
    addTranscript()
    )
    
  }
}

Masterfile <- do.call(rbind, lapply(resLFC_list, get)) %>%     
  mutate(direction = ifelse(log2FoldChange > 0, "up", ifelse(log2FoldChange < 0, "down", NA))) 

Masterfile_padj <- Masterfile %>% subset(log2FoldChange >= 1 & log2FoldChange > 0 | log2FoldChange <= -1 & log2FoldChange < 0) %>% subset(padj < 0.01)
```

# Gene body coverage (Figure S1)
For generating the gene body coverage statistics see: RESeQCoverage.sh
```{r}
coverage_files <- list.files(here::here("data/gene_body_coverage"), pattern = "txt")
coverage <- list()

for (i in 1:length(coverage_files)) {
  dat <- read.delim(paste0(here::here("data/gene_body_coverage/"), coverage_files[i]), 
                           sep = "\t", header = T) 
  dat <- dat %>% 
    melt(id.vars = "Percentile")
  dat$value <- scales::rescale(dat$value)
  
  coverage <- rlist::list.append(
    coverage,
    dat
  )
}

coverage <- do.call(rbind, coverage)

coverage$variable <- as.character(coverage$variable)
coverage$percentile <- 0

for (i in 1:nrow(coverage)) {
  coverage$percentile[i] <- as.numeric(substr(as.character(coverage$variable[i]), 2, nchar(as.character(coverage$variable[i]))))
}
coverage$percentile <- as.numeric(coverage$percentile)
colnames(coverage) <- c("sample", "id", "value", "percentile")


coverage %>% 
  ggplot(aes(x = percentile, y = value, color = sample)) +
  geom_line(size = 0.2) +
  labs(x = "Gene body percentile (5'->3')",
       y = "Coverage",
       color = "Sample") +
  scale_color_viridis_d() +
  theme_bw() +
  theme(legend.position="bottom",
        text = element_text(size = 16),
        legend.text = element_text(size = 10))
```

# Gene expression correlation (Figure S2A-D)
```{r}
directory <- here::here("data/seq_data")
A1 <- read.csv(paste0(directory, "/A1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
A1_repl <- read.csv(paste0(directory, "/A1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
A6 <- read.csv(paste0(directory, "/A6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B1 <- read.csv(paste0(directory, "/B1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B1_repl <- read.csv(paste0(directory, "/B1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B6 <- read.csv(paste0(directory, "/B6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C1 <- read.csv(paste0(directory, "/C1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C1_repl <- read.csv(paste0(directory, "/C1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C6 <- read.csv(paste0(directory, "/C6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D1 <- read.csv(paste0(directory, "/D1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D1_repl <- read.csv(paste0(directory, "/D1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D6 <- read.csv(paste0(directory, "/D6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E1 <- read.csv(paste0(directory, "/E1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E1_repl <- read.csv(paste0(directory, "/E1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E6 <- read.csv(paste0(directory, "/E6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")

colnames(A1) <- c("Transcript ID", "A1")
colnames(A1_repl) <- c("Transcript ID", "A1_repl")
colnames(A6) <- c("Transcript ID", "A6")
colnames(B1) <- c("Transcript ID", "B1")
colnames(B1_repl) <- c("Transcript ID", "B1_repl")
colnames(B6) <- c("Transcript ID", "B6")
colnames(C1) <- c("Transcript ID", "C1")
colnames(C1_repl) <- c("Transcript ID", "C1_repl")
colnames(C6) <- c("Transcript ID", "C6")
colnames(D1) <- c("Transcript ID", "D1")
colnames(D1_repl) <- c("Transcript ID", "D1_repl")
colnames(D6) <- c("Transcript ID", "D6")
colnames(E1) <- c("Transcript ID", "E1")
colnames(E1_repl) <- c("Transcript ID", "E1_repl")
colnames(E6) <- c("Transcript ID", "E6")


merge_A <- inner_join(A1, A1_repl, by="Transcript ID") 
merge_A <- inner_join(merge_A, A6, by="Transcript ID") 

merge_B <- inner_join(B1, B1_repl, by="Transcript ID") 
merge_B <- inner_join(merge_B, B6, by="Transcript ID") 

merge_C <- inner_join(C1, C1_repl, by="Transcript ID") 
merge_C <- inner_join(merge_C, C6, by="Transcript ID") 

merge_D <- inner_join(D1, D1_repl, by="Transcript ID") 
merge_D <- inner_join(merge_D, D6, by="Transcript ID") 

merge_E <- inner_join(E1, E1_repl, by="Transcript ID") 
merge_E <- inner_join(merge_E, E6, by="Transcript ID") 

Merge_4hr <- inner_join(merge_A, merge_B, by="Transcript ID")
Merge_4hr <- inner_join(Merge_4hr, merge_C, by="Transcript ID")
Merge_4hr <- inner_join(Merge_4hr, merge_D, by="Transcript ID")
Merge_4hr <- inner_join(Merge_4hr, merge_E, by="Transcript ID")
Merge_4hr

rownames(Merge_4hr) <- Merge_4hr$`Transcript ID`
Merge_4hr <- dplyr::select(Merge_4hr, -`Transcript ID`)

Merge_4hr_corr <- cor(Merge_4hr)


Merge_4hr_corr_plot <- corrplot(Merge_4hr_corr, 
                                method="number", type="upper", tl.col="black", number.cex=0.5, 
                                col= rev(brewer.pal(n=8, name="RdBu")))

Merge_4hr_corr_plot

corrplot(Merge_4hr_corr, 
         method="number", type="upper", tl.col="black", number.cex=0.5, 
         col= rev(brewer.pal(n=8, name="RdBu")))
```

```{r}
#-- Technical replicate correlation matrix plot

A1.1 <- read.csv(paste0(directory, "/A1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
A1.2 <- read.csv(paste0(directory, "/A1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B1.1 <- read.csv(paste0(directory, "/B1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B1.2 <- read.csv(paste0(directory, "/B1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C1.1 <- read.csv(paste0(directory, "/C1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C1.2 <- read.csv(paste0(directory, "/C1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D1.1 <- read.csv(paste0(directory, "/D1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D1.2 <- read.csv(paste0(directory, "/D1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E1.1 <- read.csv(paste0(directory, "/E1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E1.2 <- read.csv(paste0(directory, "/E1_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")


colnames(A1.1) <- c("Transcript ID", "A1.1")
colnames(A1.2) <- c("Transcript ID", "A1.2")
colnames(B1.1) <- c("Transcript ID", "B1.1")
colnames(B1.2) <- c("Transcript ID", "B1.2")
colnames(C1.1) <- c("Transcript ID", "C1.1")
colnames(C1.2) <- c("Transcript ID", "C1.2")
colnames(D1.1) <- c("Transcript ID", "D1.1")
colnames(D1.2) <- c("Transcript ID", "D1.2")
colnames(E1.1) <- c("Transcript ID", "E1.1")
colnames(E1.2) <- c("Transcript ID", "E1.2")


merge_tech_A <- inner_join(A1.1, A1.2, by="Transcript ID") 
merge_tech_B <- inner_join(B1.1, B1.2, by="Transcript ID") 
merge_tech_C <- inner_join(C1.1, C1.2, by="Transcript ID") 
merge_tech_D <- inner_join(D1.1, D1.2, by="Transcript ID") 
merge_tech_E <- inner_join(E1.1, E1.2, by="Transcript ID") 


Merge_tech_4hr <- inner_join(merge_tech_A, merge_tech_B, by="Transcript ID")
Merge_tech_4hr <- inner_join(Merge_tech_4hr, merge_tech_C, by="Transcript ID")
Merge_tech_4hr <- inner_join(Merge_tech_4hr, merge_tech_D, by="Transcript ID")
Merge_tech_4hr <- inner_join(Merge_tech_4hr, merge_tech_E, by="Transcript ID")
#Merge_tech_4hr

rownames(Merge_tech_4hr) <- Merge_tech_4hr$`Transcript ID`
Merge_tech_4hr <- dplyr::select(Merge_tech_4hr, -`Transcript ID`)

Merge_tech_4hr_corr <- cor(Merge_tech_4hr)


corrplot(Merge_tech_4hr_corr, method="number", type="upper", tl.col="black", number.cex=0.5, col= rev(brewer.pal(n=8, name="RdBu")))


#-- Biological replicate correlation matrix plot

A1.1 <- read.csv(paste0(directory, "/A1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
A1.3 <- read.csv(paste0(directory, "/A6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B1.1 <- read.csv(paste0(directory, "/B1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
B1.3 <- read.csv(paste0(directory, "/B6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C1.1 <- read.csv(paste0(directory, "/C1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
C1.3 <- read.csv(paste0(directory, "/C6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D1.1 <- read.csv(paste0(directory, "/D1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
D1.3 <- read.csv(paste0(directory, "/D6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E1.1 <- read.csv(paste0(directory, "/E1_RPMI_4hr_R1.counts.txt"), header = FALSE, sep="\t")
E1.3 <- read.csv(paste0(directory, "/E6_RPMI_repl_4hr_R1.counts.txt"), header = FALSE, sep="\t")


colnames(A1.1) <- c("Transcript ID", "A1.1")
colnames(A1.3) <- c("Transcript ID", "A1.3")
colnames(B1.1) <- c("Transcript ID", "B1.1")
colnames(B1.3) <- c("Transcript ID", "B1.3")
colnames(C1.1) <- c("Transcript ID", "C1.1")
colnames(C1.3) <- c("Transcript ID", "C1.3")
colnames(D1.1) <- c("Transcript ID", "D1.1")
colnames(D1.3) <- c("Transcript ID", "D1.3")
colnames(E1.1) <- c("Transcript ID", "E1.1")
colnames(E1.3) <- c("Transcript ID", "E1.3")

merge_biol_A <- inner_join(A1.1, A1.3, by="Transcript ID") 
merge_biol_B <- inner_join(B1.1, B1.3, by="Transcript ID") 
merge_biol_C <- inner_join(C1.1, C1.3, by="Transcript ID") 
merge_biol_D <- inner_join(D1.1, D1.3, by="Transcript ID") 
merge_biol_E <- inner_join(E1.1, E1.3, by="Transcript ID") 

Merge_biol_4hr <- inner_join(merge_biol_A, merge_biol_B, by="Transcript ID")
Merge_biol_4hr <- inner_join(Merge_biol_4hr, merge_biol_C, by="Transcript ID")
Merge_biol_4hr <- inner_join(Merge_biol_4hr, merge_biol_D, by="Transcript ID")
Merge_biol_4hr <- inner_join(Merge_biol_4hr, merge_biol_E, by="Transcript ID")
#Merge_biol_4hr

rownames(Merge_biol_4hr) <- Merge_biol_4hr$`Transcript ID`
Merge_biol_4hr <- dplyr::select(Merge_biol_4hr, -`Transcript ID`)

Merge_biol_4hr_corr <-cor(Merge_biol_4hr)


corrplot(Merge_biol_4hr_corr, method="number", type="lower", tl.col="black", number.cex=0.5, col= rev(brewer.pal(n=8, name="RdBu")))

corrplot(Merge_biol_4hr_corr, 
         method="number", type="lower", tl.col="black", number.cex=0.5, col= rev(brewer.pal(n=8, name="RdBu")))
```

Correlation matrix, separate for biological and technical replicates- 24 hours
```{r}
#-- Technical replicate correlation matrix plot

A7.1 <- read.csv(paste0(directory, "/A7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
A7.2 <- read.csv(paste0(directory, "/A7_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
B7.1 <- read.csv(paste0(directory, "/B7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
B7.2 <- read.csv(paste0(directory, "/B7_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
C7.1 <- read.csv(paste0(directory, "/C7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
C7.2 <- read.csv(paste0(directory, "/C7_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
D7.1 <- read.csv(paste0(directory, "/D7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
D7.2 <- read.csv(paste0(directory, "/D7_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
E7.1 <- read.csv(paste0(directory, "/E7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
E7.2 <- read.csv(paste0(directory, "/E7_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")


colnames(A7.1) <- c("Transcript ID", "A7.1")
colnames(A7.2) <- c("Transcript ID", "A7.2")
colnames(B7.1) <- c("Transcript ID", "B7.1")
colnames(B7.2) <- c("Transcript ID", "B7.2")
colnames(C7.1) <- c("Transcript ID", "C7.1")
colnames(C7.2) <- c("Transcript ID", "C7.2")
colnames(D7.1) <- c("Transcript ID", "D7.1")
colnames(D7.2) <- c("Transcript ID", "D7.2")
colnames(E7.1) <- c("Transcript ID", "E7.1")
colnames(E7.2) <- c("Transcript ID", "E7.2")


merge_tech_A <- inner_join(A7.1, A7.2, by="Transcript ID") 
merge_tech_B <- inner_join(B7.1, B7.2, by="Transcript ID") 
merge_tech_C <- inner_join(C7.1, C7.2, by="Transcript ID") 
merge_tech_D <- inner_join(D7.1, D7.2, by="Transcript ID") 
merge_tech_E <- inner_join(E7.1, E7.2, by="Transcript ID") 


Merge_tech_24hr <- inner_join(merge_tech_A, merge_tech_B, by="Transcript ID")
Merge_tech_24hr <- inner_join(Merge_tech_24hr, merge_tech_C, by="Transcript ID")
Merge_tech_24hr <- inner_join(Merge_tech_24hr, merge_tech_D, by="Transcript ID")
Merge_tech_24hr <- inner_join(Merge_tech_24hr, merge_tech_E, by="Transcript ID")
#Merge_tech_24hr

rownames(Merge_tech_24hr) <- Merge_tech_24hr$`Transcript ID`
Merge_tech_24hr <- dplyr::select(Merge_tech_24hr, -`Transcript ID`)

Merge_tech_24hr_corr <-cor(Merge_tech_24hr)


corrplot(Merge_tech_24hr_corr, method="number", type="upper", tl.col="black", number.cex=0.7, col= rev(brewer.pal(n=8, name="RdBu")))

corrplot(Merge_tech_24hr_corr, 
         method="number", type="upper", tl.col="black", number.cex=0.7, col= rev(brewer.pal(n=8, name="RdBu")))

#-- Biological replicate correlation matrix plot

A7.1 <- read.csv(paste0(directory, "/A7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
A7.3 <- read.csv(paste0(directory, "/A12_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
B7.1 <- read.csv(paste0(directory, "/B7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
B7.3 <- read.csv(paste0(directory, "/B12_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
C7.1 <- read.csv(paste0(directory, "/C7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
C7.3 <- read.csv(paste0(directory, "/C12_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
D7.1 <- read.csv(paste0(directory, "/D7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
D7.3 <- read.csv(paste0(directory, "/D12_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")
E7.1 <- read.csv(paste0(directory, "/E7_RPMI_24hr_R1.counts.txt"), header = FALSE, sep="\t")
E7.3 <- read.csv(paste0(directory, "/E12_RPMI_repl_24hr_R1.counts.txt"), header = FALSE, sep="\t")


colnames(A7.1) <- c("Transcript ID", "A7.1")
colnames(A7.3) <- c("Transcript ID", "A7.3")
colnames(B7.1) <- c("Transcript ID", "B7.1")
colnames(B7.3) <- c("Transcript ID", "B7.3")
colnames(C7.1) <- c("Transcript ID", "C7.1")
colnames(C7.3) <- c("Transcript ID", "C7.3")
colnames(D7.1) <- c("Transcript ID", "D7.1")
colnames(D7.3) <- c("Transcript ID", "D7.3")
colnames(E7.1) <- c("Transcript ID", "E7.1")
colnames(E7.3) <- c("Transcript ID", "E7.3")

merge_biol_A <- inner_join(A7.1, A7.3, by="Transcript ID") 
merge_biol_B <- inner_join(B7.1, B7.3, by="Transcript ID") 
merge_biol_C <- inner_join(C7.1, C7.3, by="Transcript ID") 
merge_biol_D <- inner_join(D7.1, D7.3, by="Transcript ID") 
merge_biol_E <- inner_join(E7.1, E7.3, by="Transcript ID") 

Merge_biol_24hr <- inner_join(merge_biol_A, merge_biol_B, by="Transcript ID")
Merge_biol_24hr <- inner_join(Merge_biol_24hr, merge_biol_C, by="Transcript ID")
Merge_biol_24hr <- inner_join(Merge_biol_24hr, merge_biol_D, by="Transcript ID")
Merge_biol_24hr <- inner_join(Merge_biol_24hr, merge_biol_E, by="Transcript ID")
#Merge_biol_24hr

rownames(Merge_biol_24hr) <- Merge_biol_24hr$`Transcript ID`
Merge_biol_24hr <- dplyr::select(Merge_biol_24hr, -`Transcript ID`)

Merge_biol_24hr_corr <-cor(Merge_biol_24hr)


corrplot(Merge_biol_24hr_corr, method="number", type="lower", tl.col="black", number.cex=0.7, col= rev(brewer.pal(n=8, name="RdBu")))

corrplot(Merge_biol_24hr_corr, 
         method="number", type="lower", tl.col="black", number.cex=0.7, col= rev(brewer.pal(n=8, name="RdBu")))
```

# Principal component analysis (Figure 1B and S10A-B)
Generating a combined dds
```{r}
sampleFilesAll <- grep("*4hr_R1", list.files(here("data/seq_data")), value = TRUE)

sampleTableAll <- data.frame(
  sampleName = (metadat[grep("*4hr", metadat$id), ]$id), 
  fileName = (gsub("$", ".counts.txt", (metadat[grep("*4hr", metadat$id), ]$id))), 
  stimulus = (metadat[grep("*4hr", metadat$id), ]$stimulus), 
  donor = (metadat[grep("*4hr", metadat$id), ]$donor), 
  time_point = (metadat[grep("*4hr", metadat$id), ]$time)
  )

ddsHTSeq_All <- DESeqDataSetFromHTSeqCount(
  sampleTable = sampleTableAll, 
  directory = here::here("data/seq_data"), 
  design = ~ donor + time_point + stimulus
  )

keep_All <- rowSums(counts(ddsHTSeq_All)) >= 10
ddsHTSeq_All <- ddsHTSeq_All[keep_All, ]

ddsHTSeq_All$stimulus <- relevel(ddsHTSeq_All$stimulus, ref="RPMI")
#ddsAll <- DESeq(ddsHTSeq_All)

#This is the data that will be used for the 4&24hrs PCAs
ntd_All <- normTransform(ddsHTSeq_All)
colData(ntd_All)

#And data for the 4 and 24 hr situations separately
ddsHTSeq_4hrs <- readRDS(here("output/ddsHTSeq_4hrs.rda"))
ddsHTSeq_24hrs <- readRDS(here("output/ddsHTSeq_24hrs.rda"))
```

A theme for the PCA plots
```{r}
theme <- theme(panel.background = element_blank(),
               panel.border=element_rect(fill=NA),
               panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(),
               strip.background=element_blank(),
               axis.ticks=element_line(colour="black"),
               plot.margin=unit(c(1,1,1,1),"line"),
               #axis.text.x = element_text(vjust=0, size = 16, hjust = 0), 
               #axis.text.y = element_text(vjust=1, size = 16, hjust=0),
               text=element_text(size=24),
               legend.position = "top",
               legend.text = element_text(size=24)
               )
```

Main PCA (Figure 1B)
```{r}
pcaData <- plotPCA(ntd_All, 
                   intgroup=c("stimulus", "time_point"), 
                   returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))

PCA_4_24hrs_2 <-
  ggplot(pcaData, aes(PC1, PC2, shape = time_point, color = stimulus)) +
  geom_point(aes(), size=4) + 
  scale_color_viridis_d(limits = c("RPMI", "C_Albicans", "LPS",  "PolyIC", "S_Aureus")) +
  #geom_mark_ellipse(color = "black")+
  ggtitle("PCA 4 and 24 hours") +
  scale_x_continuous(limits = c(-50,50), expand=c(0,0))+
  scale_y_continuous(limits = c(-50,50), expand=c(0,0))+
  scale_shape_discrete(name = "Time-point")+
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  theme

PCA_4_24hrs_2
```

PCA 4 hours (Figure S10A)
```{r}
ntd_4hrs <- normTransform(ddsHTSeq_4hrs)
colData(ntd_4hrs)

pcaData <- plotPCA(ntd_4hrs, 
                   intgroup=c("group"), 
                   returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))

pcaData <- pcaData %>% mutate(donor = substr(name, 1, 1))


PCA_4hrs <- pcaData %>% 
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = group, shape = donor), size=4) + 
  scale_color_viridis_d(limits = c("RPMI", "C_Albicans", "LPS",  "PolyIC", "S_Aureus")) +
  ggtitle("PCA 4 hours") +
  scale_x_continuous(limits = c(-50,50), expand=c(0,0))+
  scale_y_continuous(limits = c(-50,50), expand=c(0,0))+
  scale_shape_discrete(name = "Donor")+
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  guides(color = guide_legend(nrow=2, byrow=TRUE),
         shape = guide_legend(nrow=2, byrow=TRUE)) +
  theme +
  theme(legend.text=element_text(size=14))
```

PCA 24 hrs (Figure S10B)
```{r}
ntd_24hrs<- normTransform(ddsHTSeq_24hrs)
colData(ntd_24hrs)

pcaData <- plotPCA(ntd_24hrs, 
                   intgroup=c("group"), 
                   returnData=TRUE
                   )
percentVar <- round(100 * attr(pcaData, "percentVar"))

pcaData <- pcaData %>% mutate(donor = substr(name, 1, 1))

PCA_24hrs <- pcaData %>% 
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = group, shape = donor), size=4) + 
  scale_color_viridis_d(limits = c("RPMI", "C_Albicans", "LPS",  "PolyIC", "S_Aureus")) +
  ggtitle("PCA 24 hours") +
  scale_x_continuous(limits = c(-60,50), expand=c(0,0))+
  scale_y_continuous(limits = c(-50,50), expand=c(0,0))+
  scale_shape_discrete(name = "Donor")+
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    guides(color = guide_legend(nrow=2, byrow=TRUE),
         shape = guide_legend(nrow=2, byrow=TRUE)) +
  theme +
  theme(legend.text=element_text(size=14))
```

# Hierarchical clustering (Figure 1C and S3A-B)
Preparation (Figure S3A-B)
```{r}
dds_4hr <- readRDS(here::here("output/dds_4hr.rda"))
dds_24hr <- readRDS(here::here("output/dds_24hr.rda"))
  
set.seed(24)
dds_all <- full_join(counts(dds_4hr) %>% as.data.frame() %>% tibble::rownames_to_column(), 
                     counts(dds_24hr) %>% as.data.frame() %>% tibble::rownames_to_column(),
                     by = "rowname") %>%
  tibble::remove_rownames() %>% 
  tibble::column_to_rownames("rowname") %>% 
  na.omit() 

t_t <- as.matrix(dds_all) 

t_t <- DESeq2::vst(t_t)
t_t <- t_t %>% 
  as.data.frame()

top500 <- head(order(rowVars(as.matrix(t_t)), decreasing = TRUE), 500)
t_t <- as.matrix(t_t[top500,])
 
degs_kmeans_genes_top500_rlog <- fviz_nbclust(t_t, kmeans, method = "wss") + 
  theme(text = element_text(size = 20)) + ggtitle("Optimal number of clusters: Genes")

degs_kmeans_samples_top500_rlog <- fviz_nbclust(t(t_t), kmeans, method = "wss") + 
  theme(text = element_text(size = 20)) + ggtitle("Optimal number of clusters: Samples")
```

Clustering results (Figure 1C)
```{r}
annotation_df <- colnames(t_t) %>% as.data.frame() 
colnames(annotation_df) <- "ID"
annotation_df <- annotation_df %>% 
  mutate(`Time-point` = str_split(ID, "_") %>% purrr::map_chr(.,3)) %>% 
  mutate(`Time-point` = ifelse(`Time-point` == "repl", str_split(ID, "_") %>% purrr::map_chr(.,4), `Time-point`)) %>% 
  mutate(`Condition` = str_split(ID, "_") %>% purrr::map_chr(.,2)) %>% 
  mutate(Individual = substr(ID, 1, 1)) %>% 
  tibble::column_to_rownames("ID") 

individual_colors <- viridis::inferno(n=5)
names(individual_colors) <- unique(annotation_df$Individual)

condition_colors <- c("#3e1a51", "#478f8d", "#41528a", "#7dbc6b", "#f7ea4b")
names(condition_colors) <- c("RPMI", "LPS", "Calbicans", "PolyIC", "Saureus")

time_colors <- c("#d9dcd9", "#bdbccf")
names(time_colors) <- unique(annotation_df$`Time-point`)


ann_colors = list(Individual = individual_colors, `Time-point` = time_colors, Condition = condition_colors)


heat <- pheatmap::pheatmap(t_t, cutree_rows = 6, cutree_cols = 6, show_rownames = F, fontsize = 14, cellheight=1.1,cellwidth=16, annotation_col = annotation_df, annotation_colors = ann_colors) 


#Output for supplemental
cluster_df_gene <- sort(cutree(heat$tree_row, k=6)) %>% as.data.frame() %>% tibble::rownames_to_column("gene")
colnames(cluster_df_gene) <- c("gene", "cluster")

cluster_df_sample <- sort(cutree(heat$tree_col, k=6)) %>% as.data.frame() %>% tibble::rownames_to_column("sample")
colnames(cluster_df_sample) <- c("sample", "cluster")
arrange(cluster_df_sample, sample)

t_out <- t_t %>% as.data.frame() %>% tibble::rownames_to_column("gene")
t_out <- full_join(t_out,
                   cluster_df_gene) %>% 
  tibble::column_to_rownames("gene")
```

# Gene set co-regulation (Figuse S11A-B)
Getting the data into shape
```{r}
dds_all <- full_join(counts(dds_4hr) %>% as.data.frame() %>% tibble::rownames_to_column(), 
                     counts(dds_24hr) %>% as.data.frame() %>% tibble::rownames_to_column(),
                     by = "rowname") %>% tibble::column_to_rownames("rowname") %>% mutate_if(is.character, as.numeric) 

dds_all <- dds_all[!rowSums(dds_all < 10), ]
cols <- as.data.frame(list(colnames(dds_all)), col.names = "id") %>% 
  mutate(individual = substr(id, 1, 1)) %>% 
  mutate(condition = str_split(id, "_", simplify = TRUE)[, 2]) %>% 
  mutate(time = ifelse(str_split(id, "_", simplify = TRUE)[, 3] %in% c("4hr", "24hr"),
                       str_split(id, "_", simplify = TRUE)[, 3],
                       str_split(id, "_", simplify = TRUE)[, 4])) %>% 
  mutate(cond_time = paste0(condition, "_", time))

order <- c(
  subset(cols, condition == "RPMI" & time == "4hr")$id,
  subset(cols, condition == "LPS" & time == "4hr")$id,
  subset(cols, condition == "Saureus" & time == "4hr")$id,
  subset(cols, condition == "Calbicans" & time == "4hr")$id,
  subset(cols, condition == "PolyIC" & time == "4hr")$id,
  
  subset(cols, condition == "RPMI" & time == "24hr")$id,
  subset(cols, condition == "LPS" & time == "24hr")$id,
  subset(cols, condition == "Saureus" & time == "24hr")$id,
  subset(cols, condition == "Calbicans" & time == "24hr")$id,
  subset(cols, condition == "PolyIC" & time == "24hr")$id
)

dds_all <- dds_all[, order]

dds_norm <- limma::normalizeBetweenArrays(log2(dds_all), method="quantile")
dds_norm <- dds_norm[order(rowMeans(dds_norm), decreasing=TRUE), ]
dds_norm <- dds_norm[!apply(dds_norm, 1, function(row) all(is.na(row))), ]
```

Getting pathways and running geseca
```{r}
pathways <- msigdbr("Homo sapiens", category="C5") %>% subset(gs_subcat == "GO:BP")
pathways <- split(as.character(pathways$ensembl_gene), pathways$gs_name)

set.seed(1)
gesecaRes <- geseca(
  pathways = pathways,
  E = dds_norm,
  minSize = 100, 
  maxSize = 500
)
gesecaRes_signif <- gesecaRes %>% subset(padj < 0.01)
```

Plotting the output (Figure S11A)
```{r}
plotGesecaTable(gesecaRes[1:20,], pathways, E = dds_norm)
```

Plotting examples (Figure S11B)
```{r}
plotCoregulationProfile(pathway=pathways[["GOBP_CYTOPLASMIC_TRANSLATION"]], 
                        E=dds_norm, titles = order, conditions = sapply(order, function(t) str_split(t, "_") %>% purrr::map_chr(.,2)))

plotCoregulationProfile(pathway=pathways[["GOBP_DEFENSE_RESPONSE_TO_OTHER_ORGANISM"]], 
                        E=dds_norm, titles = order, conditions = sapply(order, function(t) str_split(t, "_") %>% purrr::map_chr(.,2)))

plotCoregulationProfile(pathway=pathways[["GOBP_RESPONSE_TO_VIRUS"]], 
                        E=dds_norm, titles = order, conditions = sapply(order, function(t) str_split(t, "_") %>% purrr::map_chr(.,2)))

plotCoregulationProfile(pathway=pathways[["GOBP_INNATE_IMMUNE_RESPONSE"]], 
                        E=dds_norm, titles = order, conditions = sapply(order, function(t) str_split(t, "_") %>% purrr::map_chr(.,2)))
```

# Gene set enrichment analysis (Figure 2A-G and S12A-E)
Generate GSEA
```{r}
#C5 GO:BP available from GSEA: https://www.gsea-msigdb.org/gsea/msigdb/collections.jsp
pathways <- fgsea::gmtPathways(here::here("data/c5.go.bp.v7.5.symbols.gmt"))
pathname <- "GO_BP"

time <- "4hr"

LPS_4 <- fgseaFromdds(dds_4hr, test = "LPS", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500,
                      name = paste(pathname, "pathways enriched in LPS vs. RPMI at 4hrs"))

PolyIC_4 <- fgseaFromdds(dds_4hr, test = "PolyIC", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                         name = paste(pathname, "pathways enriched in PolyIC vs. RPMI  at 4hrs"))

CAlb_4 <- fgseaFromdds(dds_4hr, test = "C_Albicans", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                       name = paste(pathname, "pathways enriched in C. Albicans vs. RPMI at 4hrs"))

SAur_4 <- fgseaFromdds(dds_4hr, test = "S_Aureus", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                       name = paste(pathname, "pathways enriched in S. Aureus vs. RPMI at 4hrs"))

time <- "24hr"
LPS_24 <- fgseaFromdds(dds_24hr, test = "LPS", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                       name = paste(pathname, "pathways enriched in LPS vs. RPMI at 24hrs"))

PolyIC_24 <- fgseaFromdds(dds_24hr, test = "PolyIC", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                          name = paste(pathname, "pathways enriched in PolyIC vs. RPMI  at 24hrs"))

CAlb_24 <- fgseaFromdds(dds_24hr, test = "C_Albicans", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                        name = paste(pathname, "pathways enriched in C. Albicans vs. RPMI at 24hrs"))

SAur_24 <- fgseaFromdds(dds_24hr, test = "S_Aureus", control = "RPMI", pathways = pathways, time, pathname, 
                      min = 100,
                      max = 500, 
                        name = paste(pathname, "pathways enriched in S. Aureus vs. RPMI at 24hrs"))

pathway_res_4 <- list(CAlb_4, LPS_4, PolyIC_4, SAur_4) %>% map2(names(.), ~mutate(.x, condition = .y))
pathway_res_24 <- list(CAlb_24, LPS_24, PolyIC_24, SAur_24) %>% map2(names(.), ~mutate(.x, condition = .y))
```

Number of induced pathways (Figure S12A)
```{r}
pathways_res <- rbind(do.call(rbind, pathway_res_4) %>% mutate(time = "4hr"),
                      do.call(rbind, pathway_res_24) %>% mutate(time = "24hr"))

pathways_res_padj <- pathways_res %>% subset(padj < 0.01) %>% mutate(dir = ifelse(NES >= 0, "up", "down")) %>%
  mutate(condition = ifelse(condition == "C_Albicans", glue("<i>C. albicans<i>"), 
                       ifelse(condition == "S_Aureus", glue("<i>S. aureus<i>"), 
                              ifelse(condition == "PolyIC", "Poly(I:C)",
                                     ifelse(condition == "LPS", "LPS", condition)))))

pathways_res$condition <- factor(pathways_res$condition, levels = c("LPS", "<i>C. albicans<i>", "Poly(I:C)", "<i>S. aureus<i>"))

pathways_res_padj$time <- factor(pathways_res_padj$time, levels = c("4hr", "24hr"))
pathways_res_padj$dir <- factor(pathways_res_padj$dir, levels = c("up", "down"))
pathways_res_padj_plot <- pathways_res_padj %>% 
  dplyr::count(condition, time, dir) %>% 
  tidyr::complete(condition, time, dir, fill = list(n = 0)) %>% 
  ggplot(aes(x = time, y = n, fill = condition, label = n)) +
  geom_col(position = "dodge") +
  geom_text(aes(y = n+14), size = 3, position = position_dodge(width = .9)) +
  facet_grid(~dir) +
  scale_fill_viridis_d(begin = 0.2, end = 1) +
  labs(x = "Time point", y = "GO:Biological Process (N)", fill = "Stimulus") +
  theme_classic() +
  theme(text = element_text(size = 16), legend.text = ggtext::element_markdown())
```

Examples up/down for LPS (Figure S12B-E)
```{r}
pathname <- "GOBP_NCRNA_METABOLIC_PROCESS"
pathname <- "GOBP_RIBONUCLEOPROTEIN_COMPLEX_BIOGENESIS"
pathname <- "GOBP_RRNA_METABOLIC_PROCESS"
pathname <- "GOBP_RIBOSOME_BIOGENESIS"

ggarrange(
  getDataForGSEAPlot(dds_4hr, test = "LPS", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "4hr"),
  getDataForGSEAPlot(dds_4hr, test = "C_Albicans", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "4hr"),
  getDataForGSEAPlot(dds_4hr, test = "S_Aureus", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "4hr"),
  getDataForGSEAPlot(dds_4hr, test = "PolyIC", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "4hr"),
  
  getDataForGSEAPlot(dds_24hr, test = "LPS", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "24hr"),
  getDataForGSEAPlot(dds_24hr, test = "C_Albicans", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "24hr"),
  getDataForGSEAPlot(dds_24hr, test = "S_Aureus", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "24hr"),
  getDataForGSEAPlot(dds_24hr, test = "PolyIC", control = "RPMI", pathways = pathways, pathwayname = pathname, time = "24hr"),
    ncol = 4, nrow = 2,
  labels = c("LPS_4hr", "CAlb_4hr", "SAur_4hr", "PolyIC_4hr", "LPS_24hr", "CAlb_24hr", "SAur_24hr", "PolyIC_24hr")
)
```

Overlapping pathway sets (Figure 2A-B)
```{r}
names(pathway_res_4) <- c("C_Albicans", "LPS", "PolyIC", "S_Aureus")
names(pathway_res_24) <- c("C_Albicans", "LPS", "PolyIC", "S_Aureus")

getPaths <- function(t) {
  out <- subset(t, padj < 0.01)
  return(out)
}

getUp <- function(t) {
  out <- subset(t, NES > 0)$pathway
  return(out)
}

getDown <- function(t) {
  out <- subset(t, NES < 0)$pathway
  return(out)
}

names(pathway_res_4) <- pathway_res[grep("_4hr", pathway_res)]
venn::venn(list(getUp(getPaths(pathway_res_4$LPS)),
                getUp(getPaths(pathway_res_4$C_Albicans)),
                getUp(getPaths(pathway_res_4$PolyIC)),
                getUp(getPaths(pathway_res_4$S_Aureus))),
           ggplot = T,
           ilcs = 1.5,
           sncs = 1.5,
           snames = "LPS, C. albicans, Poly(I:C), S. aureus",
           ilabel = "counts",
           box = F,
           zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))

venn::venn(list(getDown(getPaths(pathway_res_4$LPS)),
                getDown(getPaths(pathway_res_4$C_Albicans)),
                getDown(getPaths(pathway_res_4$PolyIC)),
                getDown(getPaths(pathway_res_4$S_Aureus))),
           ggplot = T,
           ilcs = 1.5,
           sncs = 1.5,
           ilabel = "counts",
           snames = "LPS, C. albicans, Poly(I:C), S. aureus",
           box = F,
           zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))


names(pathway_res_24) <- pathway_res[grep("_24hr", pathway_res)]
venn::venn(list(getUp(getPaths(pathway_res_24$LPS)),
                getUp(getPaths(pathway_res_24$C_Albicans)),
                getUp(getPaths(pathway_res_24$PolyIC)),
                getUp(getPaths(pathway_res_24$S_Aureus))),
           ggplot = T,
           ilcs = 1.5,
           sncs = 1.5,
           ilabel = "counts",
           snames = "LPS, C. albicans, Poly(I:C), S. aureus",
           box = F,
           zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))

venn::venn(list(getDown(getPaths(pathway_res_24$LPS)),
                getDown(getPaths(pathway_res_24$C_Albicans)),
                getDown(getPaths(pathway_res_24$PolyIC)),
                getDown(getPaths(pathway_res_24$S_Aureus))),
           ggplot = T,
           ilcs = 1.5,
           sncs = 1.5,
           ilabel = "counts",
           snames = "LPS, C. albicans, Poly(I:C), S. aureus",
           box = F,
           zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))
```

Jaccard distances between enriched gene sets (Figure 2C)
```{r}
pathway_res_4_padj <- lapply(pathway_res_4, function(x) subset(x, padj < 0.01))
pathway_res_24_padj <- lapply(pathway_res_24, function(x) subset(x, padj < 0.01))

path_time_list <- list(
  subset(pathway_res_4_padj$C_Albicans, NES > 0)$pathway,
  subset(pathway_res_24_padj$C_Albicans, NES > 0)$pathway,
  subset(pathway_res_4_padj$C_Albicans, NES < 0)$pathway,
  subset(pathway_res_24_padj$C_Albicans, NES < 0)$pathway,
  subset(pathway_res_4_padj$LPS, NES > 0)$pathway,
  subset(pathway_res_24_padj$LPS, NES > 0)$pathway,
  subset(pathway_res_4_padj$LPS, NES < 0)$pathway,
  subset(pathway_res_24_padj$LPS, NES < 0)$pathway,
  subset(pathway_res_4_padj$S_Aureus, NES > 0)$pathway,
  subset(pathway_res_24_padj$S_Aureus, NES > 0)$pathway,
  subset(pathway_res_4_padj$S_Aureus, NES < 0)$pathway,
  subset(pathway_res_24_padj$S_Aureus, NES < 0)$pathway,
  subset(pathway_res_4_padj$PolyIC, NES > 0)$pathway,
  subset(pathway_res_24_padj$PolyIC, NES > 0)$pathway,
  subset(pathway_res_4_padj$PolyIC, NES < 0)$pathway,
  subset(pathway_res_24_padj$PolyIC, NES < 0)$pathway
)

names(path_time_list) <- c("CAlb_4_up", "CAlb_24_up", "CAlb_4_down", "CAlb_24_down",
                           "LPS_4_up", "LPS_24_up", "LPS_4_down", "LPS_24_down",
                           "SAur_4_up", "SAur_24_up", "SAur_4_down", "SAur_24_down",
                           "PolIC_4_up", "PolIC_24_up", "PolIC_4_down", "PolIC_24_down")
jaccard_path <- matrix(ncol = length(path_time_list), nrow = length(path_time_list))
       
for (i in 1:nrow(jaccard_path)) {
  for (j in 1:ncol(jaccard_path)) {
    jaccard_path[i,j] <- jaccard(path_time_list[[i]], path_time_list[[j]])
  }
}

rownames(jaccard_path) <- names(path_time_list)
colnames(jaccard_path) <- names(path_time_list)

t <- jaccard_path %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("sample_1") %>% 
  melt(id.vars = "sample_1", measure.vars = names(path_time_list)) 

custom_order <- c("CAlb_4_up", "LPS_4_up", "SAur_4_up", "PolIC_4_up", 
                  "CAlb_24_up", "LPS_24_up", "SAur_24_up", "PolIC_24_up",
                  "CAlb_4_down", "LPS_4_down", "SAur_4_down", "PolIC_4_down", 
                  "CAlb_24_down", "LPS_24_down", "SAur_24_down", "PolIC_24_down")
t$sample_1 <- factor(t$sample_1, levels = custom_order)
t$variable <- factor(t$variable, levels = custom_order)
t %>% 
  ggplot(aes(x = sample_1, y = variable, fill = value, label = round(value, 2))) +
  geom_tile() +
  geom_text() +
  scale_fill_distiller(direction = -1) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

Overlap between time points: each stimulus separate (Figure 2D)
```{r}
venn::venn(list(getUp(getPaths(pathway_res_4$LPS)), getUp(getPaths(pathway_res_24$LPS))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getUp(getPaths(pathway_res_4$C_Albicans)), getUp(getPaths(pathway_res_24$C_Albicans))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getUp(getPaths(pathway_res_4$PolyIC)), getUp(getPaths(pathway_res_24$PolyIC))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getUp(getPaths(pathway_res_4$S_Aureus)), getUp(getPaths(pathway_res_24$S_Aureus))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getDown(getPaths(pathway_res_4$LPS)), getDown(getPaths(pathway_res_24$LPS))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getDown(getPaths(pathway_res_4$C_Albicans)), getDown(getPaths(pathway_res_24$C_Albicans))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getDown(getPaths(pathway_res_4$PolyIC)), getDown(getPaths(pathway_res_24$PolyIC))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)

venn::venn(list(getDown(getPaths(pathway_res_4$S_Aureus)), getDown(getPaths(pathway_res_24$S_Aureus))
  ), ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "4 hours, 24 hours", box = F, zcolor = as.vector(viridis::viridis_pal()(2)[1:2]), ilabels = "counts"
)
```

Example pathways (Figure 2E)
```{r}
updown_LPS <- intersect(subset(pathway_res_4_padj$LPS, NES > 0)$pathway,
          subset(pathway_res_24_padj$LPS, NES < 0)$pathway)

updown_LPS_val <- rbind(
  pathway_res_4$LPS %>% mutate(time = "4hr"),
  pathway_res_24$LPS %>% mutate(time = "24hr")
) %>% 
  mutate(updown = ifelse(pathway %in% updown_LPS, T, F)) %>% 
  mutate(signif = ifelse(padj < 0.01, T, F))

updown_LPS_signif <- subset(updown_LPS_val, signif == T) %>% group_by(pathway) %>% summarize(n = n()) %>% subset(n > 1)
updown_LPS_signif$pathway %>% unique()

updown_LPS_val$time <- factor(updown_LPS_val$time, levels = c("4hr", "24hr"))
dat <- updown_LPS_val %>% 
  mutate(show = ifelse(pathway %in% updown_LPS_signif$pathway, T, F)) 

ggplot(dat, aes(x = time, y = NES, group = pathway, alpha = updown)) +
  geom_point(aes(color = interaction(show, updown))) +
  geom_line(aes(color = interaction(show, updown))) +
  geom_text(data = subset(dat, updown == T & signif == T & show == T), aes(label = pathway)) +
  scale_color_viridis_d() +
  facet_grid() +
  theme_classic()
```

Core pathways (Figure 2F)
```{r}
core_up <- venn::venn(list(
  Reduce(intersect, list(
    getUp(getPaths(pathway_res_4$LPS)),
    getUp(getPaths(pathway_res_4$C_Albicans)),
    getUp(getPaths(pathway_res_4$PolyIC)),
    getUp(getPaths(pathway_res_4$S_Aureus)))),
  Reduce(intersect, list(
    getUp(getPaths(pathway_res_24$LPS)),
    getUp(getPaths(pathway_res_24$C_Albicans)),
    getUp(getPaths(pathway_res_24$PolyIC)),
    getUp(getPaths(pathway_res_24$S_Aureus))))
  ), ilabels = "counts", ggplot = T, box = F, snames = "4 hours, 24 hours", zcolor = as.vector(viridis::viridis_pal()(2)[1:2]))

core_down <- venn::venn(list(
  Reduce(intersect, list(
    getDown(getPaths(pathway_res_4$LPS)),
    getDown(getPaths(pathway_res_4$C_Albicans)),
    getDown(getPaths(pathway_res_4$PolyIC)),
    getDown(getPaths(pathway_res_4$S_Aureus)))),
  Reduce(intersect, list(
    getDown(getPaths(pathway_res_24$LPS)),
    getDown(getPaths(pathway_res_24$C_Albicans)),
    getDown(getPaths(pathway_res_24$PolyIC)),
    getDown(getPaths(pathway_res_24$S_Aureus))))
  ), ilabels = "counts", ggplot = T, box = F, snames = "4 hours, 24 hours", zcolor = as.vector(viridis::viridis_pal()(2)[1:2]))
```

## Revigo (Figure S4 and S13)
Loading data
```{r}
#Similarity matrix generated on http://revigo.irb.hr/
revigo_table <- read_tsv(here::here("data/Revigo_BP_OnScreenTable.tsv"))

revigo_matrix <- read.delim(here::here("data/Revigo_BP_SimilarityMatrix.tsv")) %>% tibble::column_to_rownames("X") 

revigo_matrix %>% as.matrix() %>% heatmap()

get_revi_selection <- revigo_table$TermID
get_revi_selection <- stringr::str_replace(get_revi_selection, ":", ".")
colnames(revigo_matrix)[colnames(revigo_matrix) %in% get_revi_selection]
revigo_matrix_selected <- revigo_matrix[
  colnames(revigo_matrix) %in% get_revi_selection,
  colnames(revigo_matrix) %in% get_revi_selection] 

t <- stringr::str_replace(colnames(revigo_matrix_selected), "[.]", ":") 
colnames(revigo_matrix_selected) <- revigo_table[match(t, revigo_table$TermID),]$Name
rownames(revigo_matrix_selected) <- revigo_table[match(t, revigo_table$TermID),]$Name
```

Revigo clusters (Figure S4)
```{r}
fviz_nbclust(revigo_matrix_selected, kmeans, method = "wss") 
```

Revigo heatmap (Figure S13)
```{r}
#5 clusters
cluster_n <- 5

heat <- pheatmap::pheatmap(revigo_matrix_selected, 
                           cutree_rows = cluster_n, cutree_cols = cluster_n, 
                           show_rownames = T, show_colnames = F, border_color = NA) 
```

Network diagram of revigo results (Figure 2G) produced in cytoscape, see: revigo_network.cys

# Differential gene expression (Figure 3A-C, S5 and S14B-C)
Volcano plots (Figure S5)
```{r}
Masterfile$`time-point` <- factor(Masterfile$`time-point`, levels = c("4hr", "24hr"))
volcano <- Masterfile %>% 
  subset(!is.na(padj)) %>% 
  mutate(Significance = ifelse(padj < 0.01 & log2FoldChange >= 1,
                               "DEGup",
                               ifelse(padj < 0.01 & log2FoldChange <= 1,
                                      "DEGdown",
                                      "NS"))) 
#Lowest Padj
toppadj <- rbind(Masterfile %>% subset(padj < 0.01 & direction == "up") %>% group_by(`time-point`, condition, direction) %>% slice_min(order_by = padj, n = 10),
                Masterfile %>% subset(padj < 0.01 & direction == "down") %>% group_by(`time-point`, condition, direction) %>% slice_min(order_by = padj, n = 10))
ggplot(volcano) +
  theme_bw() +
  geom_point(aes(x = log2FoldChange, y = -log10(padj), color = factor(Significance)), size = 0.5) +
  geom_text_repel(data = toppadj, aes(x = log2FoldChange, y = -log10(padj), label = geneName), max.overlaps = Inf, size = 5) +
  scale_color_manual(
    values = c(
      NS = "black", 
      DEGup = viridisLite::viridis(n = 2, begin = 0.1, end = 0.9)[2], 
      DEGdown = viridisLite::viridis(n = 2, begin = 0.1, end = 0.9)[1]
    ), 
    labels = c(
      NS = "NS", 
      DEGup = paste("padj <= 0.01 & Log2FoldChange >= 1"), 
      DEGdown = paste("padj <= 0.01 & Log2FoldChange <= -1"))
  ) + 
  xlab(bquote(~Log[2]~ "fold change")) +
  facet_grid(condition~`time-point`) +
  ylab(bquote(~-Log[10]~ "adjusted" ~italic(p)~"-value")) +
  geom_vline(xintercept = c(-1, 1), 
             linetype = "longdash", 
             colour = "grey", 
             size=0.4) +
  geom_hline(yintercept = -log10(0.01), 
             linetype = "longdash", 
             colour = "grey", 
             size=0.4) +
  theme(legend.title = element_blank(),
        text = element_text(size = 24),
        legend.text = element_text(size = 24),
        legend.position = "top") +
  guides(colour = guide_legend(override.aes = list(size=1), ncol=1, nrow=4, byrow=TRUE)) 
```

PolyIC DEGs at 4hr (Figure S14B)
```{r}
polic_4hr_gprof <- gprofiler2::gost(getGenes(subset(Masterfile_padj, condition == "PolIC" & `time-point` == "4hr")))$result

plot <- polic_4hr_gprof %>% 
  subset(term_size > 100 & term_size < 500) %>% 
  subset(!source %in% c("TF", "MIRNA", "WP")) %>% 
  slice_min(p_value, n=10) %>% 
  ggplot(aes(x = -log10(p_value), y = reorder(term_name, -p_value), fill = source)) +
  geom_col() +
  #facet_grid(time~., scales = "free_y") +
  xlab(expression('-log'[10]*"P")) +
  labs(y = "Term name", fill = "Gene set source") +
  theme_classic() +
  theme(text = element_text(size = 14), legend.position = "right")
```

LPS DEGs at both time points (Figure S14C)
```{r}
LPS_compare <- rbind(
  gprofiler2::gost(query = subset(LPS, direction == "down" & time == "4hr")$geneName, 
                   organism = "hsapiens", ordered_query = FALSE, 
                   multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                   measure_underrepresentation = FALSE, evcodes = FALSE, 
                   user_threshold = 0.05, correction_method = "g_SCS", 
                   domain_scope = "annotated", custom_bg = NULL, 
                   numeric_ns = "", sources = NULL, as_short_link = FALSE)$result %>% 
    mutate(time = "4hr") %>% mutate(dir = "down"),
  
  gprofiler2::gost(query = subset(LPS, direction == "down" & time == "24hr")$geneName, 
                   organism = "hsapiens", ordered_query = FALSE, 
                   multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                   measure_underrepresentation = FALSE, evcodes = FALSE, 
                   user_threshold = 0.05, correction_method = "g_SCS", 
                   domain_scope = "annotated", custom_bg = NULL, 
                   numeric_ns = "", sources = NULL, as_short_link = FALSE)$result %>% 
    mutate(time = "24hr") %>% mutate(dir = "down")
)

plot <- LPS_compare %>% 
  subset(term_size > 100 & term_size < 500) %>% 
  mutate(time = factor(time, levels = c("4hr", "24hr"))) %>% 
  ggplot(aes(x = -log10(p_value), y = reorder(term_name, -p_value), fill = source)) +
  geom_col() +
  facet_grid(time~., scales = "free_y") +
  xlab(expression('-log'[10]*"P")) +
  labs(y = "Term name", fill = "Gene set source") +
  theme_classic() +
  theme(text = element_text(size = 12), legend.position = "right")
```

Overlap (Figure 3A-B)
```{r}
degs_4hr_up <- list(
  getGenes(subset(Masterfile_padj, condition == "LPS" & `time-point` == "4hr" & log2FoldChange >= 0)),
  getGenes(subset(Masterfile_padj, condition == "CAlb" & `time-point` == "4hr" & log2FoldChange >= 0)),
  getGenes(subset(Masterfile_padj, condition == "PolIC" & `time-point` == "4hr" & log2FoldChange >= 0)),
  getGenes(subset(Masterfile_padj, condition == "SAur" & `time-point` == "4hr" & log2FoldChange >= 0))
  )

overlap_4hr_up <- venn::venn(
  degs_4hr_up,
  ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "LPS, C. albicans, Poly(I:C), S. aureus", box = F, zcolor = as.vector(viridis::viridis_pal()(5)[2:5]), ilabels = "counts"
)

degs_4hr_down <- list(
  getGenes(subset(Masterfile_padj, condition == "LPS" & `time-point` == "4hr" & log2FoldChange <= 0)),
  getGenes(subset(Masterfile_padj, condition == "CAlb" & `time-point` == "4hr" & log2FoldChange <= 0)),
  getGenes(subset(Masterfile_padj, condition == "PolIC" & `time-point` == "4hr" & log2FoldChange <= 0)),
  getGenes(subset(Masterfile_padj, condition == "SAur" & `time-point` == "4hr" & log2FoldChange <= 0))
  )
overlap_4hr_down <- venn::venn(
  degs_4hr_down,
  ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "LPS, C. albicans, Poly(I:C), S. aureus", box = F, zcolor = as.vector(viridis::viridis_pal()(5)[2:5]), ilabels = "counts"
)

degs_24hr_up <- list(
  getGenes(subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & log2FoldChange >= 0)),
  getGenes(subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & log2FoldChange >= 0)),
  getGenes(subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & log2FoldChange >= 0)),
  getGenes(subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & log2FoldChange >= 0))
  )

overlap_24hr_up <- venn::venn(
  degs_24hr_up,
  ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "LPS, C. albicans, Poly(I:C), S. aureus", box = F, zcolor = as.vector(viridis::viridis_pal()(5)[2:5]), ilabels = "counts"
)


degs_24hr_down <- list(
  getGenes(subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & log2FoldChange <= 0)),
  getGenes(subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & log2FoldChange <= 0)),
  getGenes(subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & log2FoldChange <= 0)),
  getGenes(subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & log2FoldChange <= 0))
  )

overlap_24hr_down <- venn::venn(
  degs_24hr_down,
  ggplot = T, ilcs = 1.5, sncs = 1.5, snames = "LPS, C. albicans, Poly(I:C), S. aureus", box = F, zcolor = as.vector(viridis::viridis_pal()(5)[2:5]), ilabels = "counts"
)
```

Overlap at both time points (Figure 3C)
```{r}
LPS <- subset(Masterfile_padj, condition == "LPS")
CAlb <- subset(Masterfile_padj, condition == "CAlb")
PolIC <- subset(Masterfile_padj, condition == "PolIC")
SAur <- subset(Masterfile_padj, condition == "SAur")

#Up
venn::venn(list(Reduce(intersect, list(subset(LPS, `time-point` == "4hr" & log2FoldChange > 0)$geneName, 
                                       subset(SAur, `time-point` == "4hr" & log2FoldChange > 0)$geneName,
                                       subset(PolIC, `time-point` == "4hr" & log2FoldChange > 0)$geneName,
                                       subset(CAlb, `time-point` == "4hr" & log2FoldChange > 0)$geneName)),
                Reduce(intersect, list(subset(LPS, `time-point` == "24hr" & log2FoldChange > 0)$geneName, 
                                       subset(SAur, `time-point` == "24hr" & log2FoldChange > 0)$geneName,
                                       subset(PolIC, `time-point` == "24hr" & log2FoldChange > 0)$geneName,
                                       subset(CAlb, `time-point` == "24hr" & log2FoldChange > 0)$geneName))),
           ggplot = T,
           ilcs = 1.5,
           sncs = 1.5,
           snames = "4 hours up, 24 hours up",
           ilabels = "counts",
           box = F,
           zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))

#Down
venn::venn(list(Reduce(intersect, list(subset(LPS, `time-point` == "4hr" & log2FoldChange < 0)$geneName, 
                                       subset(SAur, `time-point` == "4hr" & log2FoldChange <  0)$geneName,
                                       subset(PolIC, `time-point` == "4hr" & log2FoldChange < 0)$geneName,
                                       subset(CAlb, `time-point` == "4hr" & log2FoldChange < 0)$geneName)),
                Reduce(intersect, list(subset(LPS, `time-point` == "24hr" & log2FoldChange < 0)$geneName, 
                                       subset(SAur, `time-point` == "24hr" & log2FoldChange < 0)$geneName,
                                       subset(PolIC, `time-point` == "24hr" & log2FoldChange < 0)$geneName,
                                       subset(CAlb, `time-point` == "24hr" & log2FoldChange < 0)$geneName))),
           ggplot = T,
           ilcs = 1.5,
           sncs = 1.5,
           ilabels = "counts",
           snames = "4 hours down, 24 hours down",
           box = F,
           zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))
```

## TLR signaling pathway (Figure S14D-F)
Gathering data. Analysis performed in cytoscape, see: tlr_plot.cys (Figure S14D-E)
```{r}
tlr_lfc_4hr <- unique(tlr_genes)
tlr_lfc_4hr <- left_join(tlr_lfc_4hr, Masterfile %>% subset(condition == "LPS" & time == "4hr") %>% 
                   dplyr::select(geneName, lfc_lps_4hr = log2FoldChange), by = "geneName")
tlr_lfc_4hr <- left_join(tlr_lfc_4hr, Masterfile %>% subset(condition == "CAlb" & time == "4hr") %>%
                   dplyr::select(geneName, lfc_calb_4hr = log2FoldChange), by = "geneName")
tlr_lfc_4hr <- left_join(tlr_lfc_4hr, Masterfile %>% subset(condition == "PolIC"  & time == "4hr") %>% 
                   dplyr::select(geneName, lfc_polic_4hr = log2FoldChange), by = "geneName")
tlr_lfc_4hr <- left_join(tlr_lfc_4hr, Masterfile %>% subset(condition == "SAur"  & time == "4hr") %>% 
                   dplyr::select(geneName, lfc_saur_4hr = log2FoldChange), by = "geneName")
tlr_lfc_4hr <- tlr_lfc_4hr %>% dplyr::select(-geneName)

tlr_lfc_24hr <- unique(tlr_genes)
tlr_lfc_24hr <- left_join(tlr_lfc_24hr, Masterfile %>% subset(condition == "LPS" & time == "24hr") %>% 
                            dplyr::select(geneName, lfc_lps_24hr = log2FoldChange), by = "geneName")
tlr_lfc_24hr <- left_join(tlr_lfc_24hr, Masterfile %>% subset(condition == "CAlb" & time == "24hr") %>%
                            dplyr::select(geneName, lfc_calb_24hr = log2FoldChange), by = "geneName")
tlr_lfc_24hr <- left_join(tlr_lfc_24hr, Masterfile %>% subset(condition == "PolIC"  & time == "24hr") %>% 
                            dplyr::select(geneName, lfc_polic_24hr = log2FoldChange), by = "geneName")
tlr_lfc_24hr <- left_join(tlr_lfc_24hr, Masterfile %>% subset(condition == "SAur"  & time == "24hr") %>% 
                            dplyr::select(geneName, lfc_saur_24hr = log2FoldChange), by = "geneName")
tlr_lfc_24hr <- tlr_lfc_24hr %>% dplyr::select(-geneName)
```

LFC distribution at 4 vs. 24hr (Figure S14F)
```{r}
padj_dist_tlr <- tlr_dat %>% reshape2::melt(id.vars = c("column"), measure.vars = c(
                                                                   "padj_lps_4hr", "padj_calb_4hr", "padj_polic_4hr", "padj_saur_4hr",
                                                                   "padj_lps_24hr", "padj_calb_24hr", "padj_polic_24hr", "padj_saur_24hr")) %>% 
  mutate(id = paste0(column, "_", str_split(variable, "_") %>% purrr::map_chr(.,2), "_", str_split(variable, "_") %>% purrr::map_chr(.,3))) %>% 
  dplyr::select(-variable)

lfc_dist_tlr <- tlr_dat %>% reshape2::melt(id.vars = c("column"), measure.vars = c("lfc_lps_4hr", "lfc_calb_4hr", "lfc_polic_4hr", "lfc_saur_4hr",
                                                                   "lfc_lps_24hr", "lfc_calb_24hr", "lfc_polic_24hr", "lfc_saur_24hr")) %>% mutate(cond = str_split(variable, "_") %>% purrr::map_chr(.,2)) %>% 
  mutate(time = str_split(variable, "_") %>% purrr::map_chr(.,3)) %>% 
  mutate(id = paste0(column, "_", str_split(variable, "_") %>% purrr::map_chr(.,2), "_", str_split(variable, "_") %>% purrr::map_chr(.,3))) %>% 
  dplyr::select(-variable)

t <- left_join(lfc_dist_tlr, padj_dist_tlr, by = "id") %>% dplyr::select(-column.y)
colnames(t) <- c("gene", "lfc", "cond", "time", "id", "padj")

lfc_dist_tlr <- t %>% mutate(signif = ifelse(padj < 0.01, T, F))


lfc_dist_tlr <- lfc_dist_tlr %>%  
  mutate(time = str_split(id, "_") %>% purrr::map_chr(.,3)) %>% 
  mutate(cond = ifelse(cond == "calb", glue("<i>C. albicans<i>"), 
                       ifelse(cond == "saur", glue("<i>S. aureus<i>"), 
                              ifelse(cond == "polic", "Poly(I:C)",
                                     ifelse(cond == "lps", "LPS", cond)))))

lfc_dist_tlr$cond <- factor(lfc_dist_tlr$cond, levels = c("LPS", "<i>C. albicans<i>", "Poly(I:C)", "<i>S. aureus<i>"))
lfc_dist_tlr$time <- factor(lfc_dist_tlr$time, levels = c("4hr", "24hr"))
lfc_dist_tlr_plot <- lfc_dist_tlr %>% 
  subset(!is.na(lfc)) %>% 
  subset(signif != F) %>% 
  mutate(lfc_cut = cut(lfc, breaks = c(-Inf, 0, 1, 2, 5, Inf))) %>% 
  group_by(cond, time, signif) %>% 
  dplyr::count(lfc_cut) %>% 
  ungroup() %>% 
  tidyr::complete(lfc_cut, cond, time, signif, fill = list(n = 0)) %>% 
  ggplot(aes(x = lfc_cut, y = n, fill = time)) +
  geom_col(position = "dodge") +
  facet_grid(~cond) +
  labs(x = "LogFoldChange", y = "TLR-associated DEGs (N)") +
  theme_classic() +
  theme(strip.text.x = ggtext::element_markdown(), legend.position = "none", text = element_text(size = 14))
```

# Gene expression validations using external datasets (Figure S15A-H)
## Harrison et al. (Figure S15A-B)
PCA (Figure S15A)
```{r}
#The harrison 2019 dataset is available at PMID: 31358949
files <- list.files(here::here("data/GSE120502_RAW"), pattern = ".txt")
matrix <- matrix(ncol = length(files), nrow = length(read.delim(paste0(here::here("data/GSE120502_RAW/"), files[1]))[,2]))

for (i in 1:length(files)) {
  matrix[, i] <- read.delim(paste0(here::here("data/GSE120502_RAW/"), files[i]))[,2]
}
colnames(matrix) <- files
rownames(matrix) <- read.delim(paste0(here::here("data/GSE120502_RAW/"), files[1]))[,1]

ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
filters <- listFilters(ensembl)
attributes <- listAttributes(ensembl)

genenames_harrison <- rownames(matrix)
geneids_harrison <- 
  biomaRt::getBM(attributes = c('hgnc_symbol', 'ensembl_gene_id', 'external_gene_name', 'external_synonym'),
                 filters = c('external_gene_name'),
                 values = genenames_harrison, 
                 mart = ensembl)

length(geneids_harrison$ensembl_gene_id) == length(rownames(matrix))

harrison_raw <- as.data.frame(matrix)
harrison_raw$geneName <- rownames(matrix)

for (i in 1:length(harrison_raw$geneName)) {
  print(i)
  
  genename <- geneids_harrison$ensembl_gene_id[geneids_harrison$external_gene_name == harrison_raw$geneName[i]] 
  harrison_raw$Transcript[i] <- ifelse(identical(genename, character(0)),
                                                 NA,
                                                 genename)
}
#For some reason, some gene names were already ensembl IDs ... these are added back in now

for (i in 1:length(harrison_raw$Transcript)) {
  if (is.na(harrison_raw$Transcript[i]) & substr(harrison_raw$geneName[i], 1, 4) == "ENSG") {
    harrison_raw$Transcript[i] <- harrison_raw$geneName[i]
  }
}

#Now, there are still some genes missing:
nrow(subset(harrison_raw, is.na(Transcript)))

#Using biomart again to gather these using the aliases
missing <- subset(harrison_raw, is.na(Transcript))$geneName
missing_harrison <- 
  biomaRt::getBM(attributes = c('hgnc_symbol', 'ensembl_gene_id', 'external_gene_name', 'external_synonym'),
                 filters = c('external_synonym'),
                 values = missing, 
                 mart = ensembl)

for (i in 1:length(harrison_raw$geneName)) {

  if (is.na(harrison_raw$Transcript[i])) {
    genename <- missing_harrison$ensembl_gene_id[missing_harrison$external_synonym == harrison_raw$geneName[i]] 
    
    harrison_raw$Transcript[i] <- ifelse(identical(genename, character(0)),
                                         NA,
                                         genename)
  }
}

#Now, there are still some genes missing:
nrow(subset(harrison_raw, is.na(Transcript)))

subset(harrison_raw, is.na(Transcript))

#For now, I'm removing these genes, maybe later we can check whether these can be added back in using either biomart or manual editing.
harrison_raw <- subset(harrison_raw, !is.na(Transcript))
n_occur <- data.frame(table(harrison_raw$Transcript))

#Removing double genes, choosing the one with the highest total reads
sumcol <- ncol(harrison_raw) -2
for (i in 1:nrow(harrison_raw)) {
  if (harrison_raw$Transcript[i] %in% n_occur[n_occur$Freq > 1,]$Var1) {
    harrison_raw$rowsum[i] <- sum(harrison_raw[i, 1:sumcol])
  } else {
    harrison_raw$rowsum[i] <- NA
  }
}
non_unique <- subset(harrison_raw, Transcript %in% n_occur[n_occur$Freq > 1,]$Var1)
non_unique <- split(non_unique, non_unique$Transcript)


getHighestExpressed <- function(t) {
  out <- t[which.max(t$rowsum),]
  return(out)
}
non_unique <- do.call(rbind, lapply(non_unique, getHighestExpressed))

unique <- subset(harrison_raw, Transcript %ni% n_occur[n_occur$Freq > 1,]$Var1)

#Putting back together
harrison_raw <- rbind(non_unique, unique)

rownames(harrison_raw) <- harrison_raw$Transcript
harrison_raw <- harrison_raw %>% dplyr::select(-geneName, -Transcript, -rowsum)
harrison_raw <- as.matrix(harrison_raw)
class(harrison_raw) <- "numeric"
harrison_raw <- as.data.frame(harrison_raw)

harrison_metadat <- matrix(ncol = 4, nrow = ncol(harrison_raw))
harrison_metadat[, 1] <- colnames(harrison_raw)
for (i in 1:ncol(harrison_raw)) {
  harrison_metadat[i, 2] <- str_split(colnames(harrison_raw)[i], "_")[[1]][2]
  harrison_metadat[i, 3] <- str_split(colnames(harrison_raw)[i], "_")[[1]][3]
  harrison_metadat[i, 4] <- ifelse(str_split(colnames(harrison_raw)[i], "_")[[1]][4] == "BATCH",
                                   paste0(str_split(colnames(harrison_raw)[i], "_")[[1]][4], "_", str_split(colnames(harrison_raw)[i], "_")[[1]][5]),
                                   str_split(colnames(harrison_raw)[i], "_")[[1]][4])
}
colnames(harrison_metadat) <- c("id", "donor", "stimulus", "batch")

dds_harrison_4hr <- DESeqDataSetFromMatrix(
  countData = harrison_raw, 
  colData = harrison_metadat, 
  design = ~ donor + stimulus
  ) %>% 
  mRNA_function %>% 
  keepFromddsHTSeq 
nrow(dds_harrison_4hr)


dds_harrison_4hr$group <- relevel(dds_harrison_4hr$group, ref = "CTL")

main <- counts(dds_4hr)
main <- cbind(rownames(main), main)
colnames(main)[1] <- "gene"

colnames(main)

harrison <- counts(dds_harrison_4hr)
harrison <- cbind(rownames(harrison), harrison)
colnames(harrison)[1] <- "gene"

harrison_master <- full_join(as.data.frame(main), as.data.frame(harrison), by = "gene")
rownames(harrison_master) <- harrison_master$gene
main <- main[, -1]
harrison_master$gene <- NULL

metadata_both <- as.data.frame(harrison_metadat)
for (i in 1:nrow(metadata_both)) {
  if(metadata_both$stimulus[i] == "CTL") {metadata_both$stimulus[i] <- "RPMI"}
  
  metadata_both$donor[i] <- paste0("Harrison_", metadata_both$donor[i])
  metadata_both$time[i] <- "4hr"
}

metadata_main <- matrix(nrow = ncol(main), ncol = 5)
for (i in 1:ncol(main)) {
  metadata_main[i, 1] <- colnames(main)[i]
  metadata_main[i, 2] <- str_split(metadata_main[i, 1], "_")[[1]][2]
  metadata_main[i, 3] <- substr(str_split(metadata_main[i, 1], "_")[[1]][1], 1, 1)
  metadata_main[i, 4] <- ifelse(str_split(metadata_main[i, 1], "_")[[1]][3] != "repl",
                                str_split(metadata_main[i, 1], "_")[[1]][3],
                                str_split(metadata_main[i, 1], "_")[[1]][4])
  metadata_main[i, 5] <- 1
}
colnames(metadata_main) <- c("id", "stimulus", "donor", "time", "batch")
metadata_main <- as.data.frame(metadata_main)
metadata_main <- subset(metadata_main, time == "4hr")

metadata_both <- rbind(metadata_main, metadata_both)
rownames(metadata_both) <- metadata_both$id


harrison_master <- drop_na(harrison_master)
harrison_master <- as.matrix(harrison_master)
class(harrison_master) <- "numeric"
harrison_master <- as.data.frame(harrison_master)

harrison_master <- 
  DESeqDataSetFromMatrix(
  countData = harrison_master, 
  colData = metadata_both, 
  design = ~ donor
  ) 

keep <- rowSums(counts(harrison_master)) >= 10
harrison_master <- harrison_master[keep, ]
harrison_master <- harrison_master %>% 
  mRNA_function

harrison_master$stimulus <- as.factor(harrison_master$stimulus)
harrison_master$stimulus <- relevel(harrison_master$stimulus, ref = "RPMI")

ntd_both <- normTransform(harrison_master)

colData(ntd_both) <- cbind(colData(ntd_both), study=c(rep("QuantSeq", 27), rep("Harrison_2019", 250)))
assay(ntd_both) <- limma::removeBatchEffect(assay(ntd_both), ntd_both$study)
pcaData <- plotPCA(ntd_both, 
                   intgroup=c("stimulus", "group"), 
                   returnData=TRUE
                   )
percentVar <- round(100 * attr(pcaData, "percentVar"))

for (i in 1:nrow(pcaData)) {
  pcaData$harrison_group[i] <- ifelse(is.na(substr(str_split(pcaData$group.1[i], "_")[[1]][2], 1, 1)),
                                      "QuantSeq",
                                      substr(str_split(pcaData$group.1[i], "_")[[1]][2], 1, 1))
}
  pcaData %>%
    mutate(donor_dataset = ifelse(group.1 %in% c("A", "B", "C", "D", "E"), "QuantSeq", "Harrison_2019")) %>% 
    ggplot(aes(x = PC1, y = PC2, shape = donor_dataset, color = stimulus)) +
    geom_point(size = 4) +
    scale_color_viridis_d() +
    ggtitle("Comparing Harrison 2019 and QuantSeq") + 
    scale_shape_discrete()+
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    theme(text = element_text(size = 24),
          aspect.ratio=1)
```

DE correlations
```{r}
t_1 <- "4hr"
t <- full_join(
  Masterfile_Harrison_2019 %>% subset(condition == "LPS") %>% subset(time == t_1),
  Masterfile %>% subset(condition == "LPS") %>% subset(time == t_1),
  by = "Transcript"
) %>% subset(!is.na(log2FoldChange.x)) %>% subset(!is.na(log2FoldChange.y)) %>% subset(!is.na(padj.x)) %>% subset(!is.na(padj.y))
t %>% 
  subset(padj.x < 0.01 & padj.y < 0.01) %>% 
  ggplot(aes(x = log2FoldChange.x, y = log2FoldChange.y)) +
  geom_point() +
  geom_smooth(method = "lm")

s <- subset(t, padj.x < 0.01 & padj.y < 0.01)
lm(s$log2FoldChange.x ~ s$log2FoldChange.y)
```

Overlap (Figure S15B)
```{r}
venn::venn(list(Masterfile_Harrison_2019$Transcript, subset(Masterfile, condition == "LPS")$Transcript), ilabels = "counts")

deg_overlap_harrison.pdf

harrison_lps_4hr <- subset(Masterfile_Harrison_2019, condition == "LPS")
quant_4hr_lps <- subset(Masterfile, condition == "LPS" & time == "4hr")

background <- intersect(Masterfile_Harrison_2019$Transcript, subset(Masterfile, condition == "LPS")$Transcript)
de_both <- Reduce(intersect, list(subset(harrison_lps_4hr, padj < 0.01)$Transcript, subset(quant_4hr_lps, padj < 0.01)$Transcript))
de_none <- background[!background %in% c(subset(harrison_lps_4hr, padj < 0.01)$Transcript, subset(quant_4hr_lps, padj < 0.01)$Transcript)]
de_only_quant <- subset(quant_4hr_lps, padj < 0.01)$Transcript[!subset(quant_4hr_lps, padj < 0.01)$Transcript %in% subset(harrison_lps_4hr, padj < 0.01)$Transcript]
de_only_har <- subset(harrison_lps_4hr, padj < 0.01)$Transcript[!subset(harrison_lps_4hr, padj < 0.01)$Transcript %in% subset(quant_4hr_lps, padj < 0.01)$Transcript]
test <- data.frame(
  c(length(de_both), length(de_only_har)),
  c(length(de_only_quant), length(de_none))
) %>% fisher.test()
```

## Corlet et al. (Figure S15C-D)
PCA (Figure S15C)
```{r}
#The corley 2019 data is available from PMID: 31827207
corley_2019 <- read.delim(here::here("validations/data/corley_2019.txt"), sep = "\t")

#Generating a metadata file
corley_2019_metadat <- matrix(ncol = 2, nrow = ncol(corley_2019)-1)
corley_2019_metadat[,1] <- colnames(corley_2019)[2:ncol(corley_2019)]
corley_2019_metadat[,2] <- c("control", "PolyIC", "control", "PolyIC", "control", "PolyIC")

colnames(corley_2019_metadat) <- c("id", "stimulus")

#Making a nice counts matrix
rownames(corley_2019) <- corley_2019[,1]
corley_2019 <- corley_2019 %>% dplyr::select(-X)

dds_corley_2019 <- DESeqDataSetFromMatrix(
  countData = corley_2019, 
  colData = corley_2019_metadat, 
  design = ~ stimulus
  ) %>% 
  keepFromddsHTSeq %>% 
  mRNA_function

dds_corley_2019$group <- relevel(dds_corley_2019$group, ref = "control")
dds_corley_2019 <- DESeq(dds_corley_2019)

resultsNames(dds_corley_2019) # lists the coefficients

res_corley <- results(dds_corley_2019)

# or to shrink log fold changes association with condition:

res_corley <- lfcShrink(dds_corley_2019, coef = "group_PolyIC_vs_control", type = "apeglm")

res_corley <- res_corley %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "PolIC") %>% tibble::rownames_to_column("Transcript")

Masterfile_corley_2019 <- res_corley %>% 
  addUpDown

for (i in 1:nrow(Masterfile_corley_2019)) {
  Masterfile_corley_2019$geneName[i] <- subset(ref_annot, Gene_id == Masterfile_corley_2019$Transcript[i])$Gene_name
}

main <- counts(ddsHTSeq_24hrs)
main <- cbind(rownames(main), main)
colnames(main)[1] <- "gene"

colnames(main)
keep <- c(1)
for (i in 1:ncol(main)) {
  if (str_split(colnames(main)[i], "_")[[1]][2] %in% c("RPMI", "PolyIC")) {
    keep <- c(keep, i)
  }
}
main <- main[, keep]

corley <- counts(dds_corley_2019)
corley <- cbind(rownames(corley), corley)
colnames(corley)[1] <- "gene"

corley_master <- full_join(as.data.frame(main), as.data.frame(corley), by = "gene")
rownames(corley_master) <- corley_master$gene
main <- main[, -1]
corley_master$gene <- NULL

metadata_both <- corley_2019_metadat
metadata_both <- as.data.frame(metadata_both) %>% mutate(donor = 1)

metadata_main <- matrix(nrow = ncol(main), ncol = 3)
for (i in 1:ncol(main)) {
  metadata_main[i, 1] <- colnames(main)[i]
  metadata_main[i, 2] <- str_split(metadata_main[i, 1], "_")[[1]][2]
  metadata_main[i, 3] <- substr(str_split(metadata_main[i, 1], "_")[[1]][1], 1, 1)
}
colnames(metadata_main) <- c("id", "stimulus", "donor")
metadata_main <- as.data.frame(metadata_main)
for (i in 1:nrow(metadata_main)) {
  if (metadata_main$stimulus[i] == "RPMI") {
    metadata_main$stimulus[i] <- "control"
  }
}

metadata_both <- rbind(metadata_main, metadata_both) %>% subset(stimulus %in% c("control", "PolyIC"))

corley_master <- drop_na(corley_master)
corley_master <- as.matrix(corley_master)
class(corley_master) <- "numeric"
corley_master <- as.data.frame(corley_master)

corley_master <- 
  DESeqDataSetFromMatrix(
  countData = corley_master, 
  colData = metadata_both, 
  design = ~ stimulus + donor
  ) 

keep <- rowSums(counts(corley_master)) >= 10
corley_master <- corley_master[keep, ]
corley_master <- corley_master %>% 
  mRNA_function

corley_master$stimulus <- relevel(corley_master$stimulus, ref = "control")

ntd_both <- normTransform(corley_master)

main <- counts(ddsHTSeq_24hrs)

colData(ntd_both) <- cbind(colData(ntd_both), study=c(rep("QuantSeq", 25), rep("Corley_2019", 6)))
assay(ntd_both) <- limma::removeBatchEffect(assay(ntd_both), ntd_both$study)
pcaData <- plotPCA(ntd_both, 
                   intgroup=c("stimulus", "group"), 
                   returnData=TRUE
                   )
percentVar <- round(100 * attr(pcaData, "percentVar"))

pcaData$donor <- pcaData$group.1
for (i in 1:nrow(pcaData)) {
  pcaData$time[i] <- ifelse(pcaData$name[i] %in% colnames(main), 
                                             ifelse(str_split(pcaData$name[i], "_")[[1]][3] %in% c("4hr", "24hr"),
                                                    str_split(pcaData$name[i], "_")[[1]][3],
                                                    str_split(pcaData$name[i], "_")[[1]][4]),
                                             pcaData$name[i])
  
  if(pcaData$time[i] %in% c("4hr", "4hrs")) {pcaData$time[i] <- "4hrs"} 
  if(pcaData$time[i] %in% c("24hr", "24hrs")) {pcaData$time[i] <- "24hrs"} 
}

  pcaData %>%
    mutate(donor_dataset = ifelse(group.1 %in% c("A", "B", "C", "D", "E"), "QuantSeq", "Corley_2019")) %>% 
    ggplot(aes(x = PC1, y = PC2, shape = donor_dataset, color = stimulus)) +
    geom_point(size = 4) +
    scale_color_viridis_d() +
    ggtitle("Comparing Corley 2019 and QuantSeq") + 
    scale_shape_discrete()+
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    theme(text = element_text(size = 24),
          aspect.ratio=1)
```

DE correlations
```{r}
t_1 <- "24hr"
t <- full_join(
  Masterfile_corley_2019 %>% subset(time == t_1),
  Masterfile %>% subset(condition == "PolIC") %>% subset(time == t_1),
  by = "Transcript"
) %>% subset(!is.na(log2FoldChange.x)) %>% subset(!is.na(log2FoldChange.y)) %>% subset(!is.na(padj.x)) %>% subset(!is.na(padj.y))
t %>% 
  subset(padj.x < 0.01 & padj.y < 0.01) %>% 
  ggplot(aes(x = log2FoldChange.x, y = log2FoldChange.y)) +
  geom_point() +
  geom_smooth(method = "lm")

s <- subset(t, padj.x < 0.01 & padj.y < 0.01)
lm(s$log2FoldChange.x ~ s$log2FoldChange.y)
```

Overlap (Figure S15D)
```{r}
venn::venn(list(Masterfile_corley_2019$Transcript, subset(Masterfile, condition == "PolIC" & time == "24hr")$Transcript), ilabels = "counts")

corley_polic_24hr <- subset(Masterfile_corley_2019)
quant_4hr_lps <- subset(Masterfile, condition == "PolIC" & time == "24hr")

background <- intersect(Masterfile_corley_2019$Transcript, subset(Masterfile, condition == "PolIC")$Transcript)
de_both <- Reduce(intersect, list(subset(corley_polic_24hr, padj < 0.01)$Transcript, subset(quant_4hr_lps, padj < 0.01)$Transcript))
de_none <- background[!background %in% c(subset(corley_polic_24hr, padj < 0.01)$Transcript, subset(quant_4hr_lps, padj < 0.01)$Transcript)]
de_only_quant <- subset(quant_4hr_lps, padj < 0.01)$Transcript[!subset(quant_4hr_lps, padj < 0.01)$Transcript %in% subset(corley_polic_24hr, padj < 0.01)$Transcript]
de_only_cor <- subset(corley_polic_24hr, padj < 0.01)$Transcript[!subset(corley_polic_24hr, padj < 0.01)$Transcript %in% subset(quant_4hr_lps, padj < 0.01)$Transcript]
test <- data.frame(
  c(length(de_both), length(de_only_cor)),
  c(length(de_only_quant), length(de_none))
) %>% fisher.test()
```

## Bruno et al. 2019 (Figure S15E-F)
PCA (Figure S15E)
```{r}
#The bruno 2019 data is available from PMID: 32839538
bruno_2019_raw <- read.delim(here::here("validations/data/bruno_2019_merged.csv"), sep = ",")
for (i in 1:nrow(bruno_2019_raw)) {
  print(i)
  rownames(bruno_2019_raw)[i] <- str_split(bruno_2019_raw$geneID[i], "[.]")[[1]][1]
}
#rownames(bruno_2019_raw) <- bruno_2019_raw$geneID
bruno_2019_raw <- bruno_2019_raw %>% dplyr::select(-geneID, -X)

for (i in 1:ncol(bruno_2019_raw)) {
  colnames(bruno_2019_raw)[i] <- str_split(colnames(bruno_2019_raw)[i], "[.]")[[1]][1]
}

bruno_2019_metadat <- matrix(ncol = 4, nrow = ncol(bruno_2019_raw))

#Generating a metadata file
meta <- bruno_2019_metadat
rownames(meta) <- colnames(bruno_2019_raw)

for (i in 1:length(rownames(meta))) {
  meta[i, 1] <- rownames(meta)[i]
  meta[i, 2] <- substr(str_split(meta[i, 1], "_")[[1]][2], 1, 1)
  meta[i, 3] <- ifelse(str_split(meta[i, 1], "_")[[1]][4] %in% c("4hrs", "24hrs"),
                       str_split(meta[i, 1], "_")[[1]][3],
                       paste0(str_split(meta[i, 1], "_")[[1]][3], "_", str_split(meta[i, 1], "_")[[1]][4]))
  meta[i, 4] <- ifelse(str_split(meta[i, 1], "_")[[1]][4] %in% c("4hrs", "24hrs"),
                       str_split(meta[i, 1], "_")[[1]][4],
                       str_split(meta[i, 1], "_")[[1]][5])
}

meta <- as.data.frame(meta)
colnames(meta) <- c("id", "donor", "stimulus", "time")
bruno_2019_metadat <- meta

#Making a nice counts matrix
bruno_2019_raw <- as.matrix(bruno_2019_raw)
class(bruno_2019_raw) <- "numeric"
bruno_2019_raw <- as.data.frame(bruno_2019_raw)

bruno_2019_metadat_4hr <- subset(bruno_2019_metadat, time == "4hrs")
bruno_2019_metadat_24hr <- subset(bruno_2019_metadat, time == "24hrs")

raw_4hr <- colnames(bruno_2019_raw) %in% bruno_2019_metadat_4hr$id %>% which()
raw_24hr <- colnames(bruno_2019_raw) %in% bruno_2019_metadat_24hr$id %>% which()

bruno_2019_raw_4hr <- bruno_2019_raw[, raw_4hr] 
bruno_2019_raw_24hr <- bruno_2019_raw[, raw_24hr] 

dds_bruno_2019_4hr <- DESeqDataSetFromMatrix(
  countData = bruno_2019_raw_4hr, 
  colData = bruno_2019_metadat_4hr, 
  design = ~ donor + stimulus
  ) %>% 
  keepFromddsHTSeq %>% 
  mRNA_function

dds_bruno_2019_4hr$group <- relevel(dds_bruno_2019_4hr$group, ref = "RPMI")
dds_bruno_2019_4hr <- DESeq(dds_bruno_2019_4hr)

resultsNames(dds_bruno_2019_4hr) # lists the coefficients

res_4hr <- results(dds_bruno_2019_4hr)

opt.cond.long <- c("group_calb_bg_vs_RPMI", 
                   "group_calb_live_vs_RPMI", 
                   "group_calb_man_vs_RPMI",
                   "group_caur_KTCbg_vs_RPMI",
                   "group_caur_KTClive_vs_RPMI",
                   "group_caur_KTCman_vs_RPMI")

# or to shrink log fold changes association with condition:
#resLFC4hrLPS <-         logFoldShrinkage(dds_4hr, opt.cond.long[1])
res_CAlb_bg_4hr <-      lfcShrink(dds_bruno_2019_4hr, coef=opt.cond.long[1], type="apeglm")
res_CAlb_live_4hr <-    lfcShrink(dds_bruno_2019_4hr, coef=opt.cond.long[2], type="apeglm")
res_CAlb_man_4hr <-     lfcShrink(dds_bruno_2019_4hr, coef=opt.cond.long[3], type="apeglm")
res_CAur_KTCbg_4hr <-   lfcShrink(dds_bruno_2019_4hr, coef=opt.cond.long[4], type="apeglm")
res_CAur_KTClive_4hr <- lfcShrink(dds_bruno_2019_4hr, coef=opt.cond.long[5], type="apeglm")
res_CAur_KTCman_4hr <-  lfcShrink(dds_bruno_2019_4hr, coef=opt.cond.long[6], type="apeglm")

res_CAlb_bg_4hr <- res_CAlb_bg_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAlb_bg") %>% tibble::rownames_to_column("Transcript")
res_CAlb_live_4hr <- res_CAlb_live_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAlb_live") %>% tibble::rownames_to_column("Transcript")
res_CAlb_man_4hr <- res_CAlb_man_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAlb_man") %>% tibble::rownames_to_column("Transcript")
res_CAur_KTCbg_4hr <- res_CAur_KTCbg_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAur_KTCb") %>% tibble::rownames_to_column("Transcript")
res_CAur_KTClive_4hr <- res_CAur_KTClive_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAur_KTClive") %>% tibble::rownames_to_column("Transcript")
res_CAur_KTCman_4hr <- res_CAur_KTCman_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAur_KTCman") %>% tibble::rownames_to_column("Transcript")

res_list_Bruno_2019_4hr <- list(res_CAlb_bg_4hr, res_CAlb_live_4hr, res_CAlb_man_4hr, 
                                res_CAur_KTCbg_4hr, res_CAur_KTClive_4hr, res_CAur_KTCman_4hr)


dds_bruno_2019_24hr <- DESeqDataSetFromMatrix(
  countData = bruno_2019_raw_24hr, 
  colData = bruno_2019_metadat_24hr, 
  design = ~ donor + stimulus
  ) %>% 
  keepFromddsHTSeq %>% 
  mRNA_function

dds_bruno_2019_24hr$group <- relevel(dds_bruno_2019_24hr$group, ref = "RPMI")
dds_bruno_2019_24hr <- DESeq(dds_bruno_2019_24hr)

resultsNames(dds_bruno_2019_24hr) # lists the coefficients

res_24hr <- results(dds_bruno_2019_24hr)

opt.cond.long <- c("group_calb_bg_vs_RPMI", 
                   "group_calb_live_vs_RPMI", 
                   "group_calb_man_vs_RPMI",
                   "group_caur_KTCbg_vs_RPMI",
                   "group_caur_KTClive_vs_RPMI",
                   "group_caur_KTCman_vs_RPMI")

# or to shrink log fold changes association with condition:
#resLFC4hrLPS <-         logFoldShrinkage(dds_4hr, opt.cond.long[1])
res_CAlb_bg_24hr <-      lfcShrink(dds_bruno_2019_24hr, coef=opt.cond.long[1], type="apeglm")
res_CAlb_live_24hr <-    lfcShrink(dds_bruno_2019_24hr, coef=opt.cond.long[2], type="apeglm")
res_CAlb_man_24hr <-     lfcShrink(dds_bruno_2019_24hr, coef=opt.cond.long[3], type="apeglm")
res_CAur_KTCbg_24hr <-   lfcShrink(dds_bruno_2019_24hr, coef=opt.cond.long[4], type="apeglm")
res_CAur_KTClive_24hr <- lfcShrink(dds_bruno_2019_24hr, coef=opt.cond.long[5], type="apeglm")
res_CAur_KTCman_24hr <-  lfcShrink(dds_bruno_2019_24hr, coef=opt.cond.long[6], type="apeglm")

res_CAlb_bg_24hr <- res_CAlb_bg_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAlb_bg") %>% tibble::rownames_to_column("Transcript")
res_CAlb_live_24hr <- res_CAlb_live_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAlb_live") %>% tibble::rownames_to_column("Transcript")
res_CAlb_man_24hr <- res_CAlb_man_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAlb_man") %>% tibble::rownames_to_column("Transcript")
res_CAur_KTCbg_24hr <- res_CAur_KTCbg_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAur_KTCbg") %>% tibble::rownames_to_column("Transcript")
res_CAur_KTClive_24hr <- res_CAur_KTClive_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAur_KTClive") %>% tibble::rownames_to_column("Transcript")
res_CAur_KTCman_24hr <- res_CAur_KTCman_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAur_KTCman") %>% tibble::rownames_to_column("Transcript")

res_list_Bruno_2019_24hr <- list(res_CAlb_bg_24hr, res_CAlb_live_24hr, res_CAlb_man_24hr, 
                                res_CAur_KTCbg_24hr, res_CAur_KTClive_24hr, res_CAur_KTCman_24hr)

Masterfile_Bruno_2019 <- rbind(
  ldply(res_list_Bruno_2019_4hr, data.frame), 
  ldply(res_list_Bruno_2019_24hr, data.frame)) %>% 
  addUpDown

for (i in 1:nrow(Masterfile_Bruno_2019)) {
  print(i)
  Masterfile_Bruno_2019$geneName[i] <- subset(ref_annot, Gene_id == Masterfile_Bruno_2019$Transcript[i])$Gene_name
}

nrow(Masterfile_Bruno_2019)

for (i in 1:nrow(Masterfile_Bruno_2019)) {
  if (Masterfile_Bruno_2019$condition[i] == "CAur_KTCb") {
    Masterfile_Bruno_2019$condition[i] <- "CAur_KTCbg"
  }
}

Masterfile_Bruno_2019_padj <- Masterfile_Bruno_2019 %>% getDEGs
nrow(Masterfile_Bruno_2019_padj)

ddsHTSeq_All <- readRDS(here::here("output/ddsHTSeq_All.rda"))
main <- counts(ddsHTSeq_All)

bruno_2019_metadat_4hr <- subset(bruno_2019_metadat, time == "4hrs")
bruno_2019_metadat_24hr <- subset(bruno_2019_metadat, time == "24hrs")

raw_4hr <- colnames(bruno_2019_raw) %in% bruno_2019_metadat_4hr$id %>% which()
raw_24hr <- colnames(bruno_2019_raw) %in% bruno_2019_metadat_24hr$id %>% which()

ntd_both
colData(ntd_both) <- cbind(colData(ntd_both), study=c(rep("QuantSeq", 52), rep("Bruno_2019", 42)))
assay(ntd_both) <- limma::removeBatchEffect(assay(ntd_both), ntd_both$study)
pcaData <- plotPCA(ntd_both, 
                   intgroup=c("stimulus", "group"), 
                   returnData=TRUE
                   )
percentVar <- round(100 * attr(pcaData, "percentVar"))

pcaData$time <- pcaData$group.1
for (i in 1:nrow(pcaData)) {
  pcaData$donor[i] <- ifelse(pcaData$name[i] %in% colnames(main), 
                                             substr(pcaData$name[i], 1, 1),
                                             paste0(substr(str_split(pcaData$name[i], "_")[[1]][2], 1, 1), "_Bruno"))
  
  if(pcaData$time[i] %in% c("4hr", "4hrs")) {pcaData$time[i] <- "4hrs"} 
  if(pcaData$time[i] %in% c("24hr", "24hrs")) {pcaData$time[i] <- "24hrs"} 
}
  
  pcaData %>%
    mutate(donor_dataset = ifelse(donor %in% c("A", "B", "C", "D", "E"), "QuantSeq", "Bruno_2019")) %>% 
    ggplot(aes(x = PC1, y = PC2, shape = interaction(time, donor_dataset), color = stimulus)) +
    geom_point(size = 4) +
    scale_color_viridis_d() +
    ggtitle("Comparing Bruno 2019 and QuantSeq") + 
    scale_shape_discrete()+
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    theme(text = element_text(size = 24),
          aspect.ratio = 1)
```

DE correlations
```{r}
t_1 <- "4hr"
t_1 <- "24hr"
stim_bruno <- unique(Masterfile_Bruno_2019$condition)[1:3]

times <- c("4hr", "24hr")
conds <- stim_bruno
for (t_1 in times) {
  for (cond in conds) {
    print(t_1)
    print(cond)
    # t_1 <- times[1]
    # cond <- conds[1]
    t <- full_join(
      Masterfile_Bruno_2019 %>% subset(condition == cond) %>% subset(time == t_1),
      Masterfile %>% subset(condition == "CAlb") %>% subset(time == t_1),
      by = "Transcript"
    ) %>% subset(!is.na(log2FoldChange.x)) %>% subset(!is.na(log2FoldChange.y)) %>% subset(!is.na(padj.x)) %>% subset(!is.na(padj.y))

    s <- subset(t, padj.x < 0.01 & padj.y < 0.01)
    test <- lm(s$log2FoldChange.x ~ s$log2FoldChange.y)

  }
}

#Live candida
#4hr overlapping DEGs (N=80): R2=0.7989, p<2.2e-16
#24hr overlapping DEGs (N=658): R2=0.07682, p=2.73e-13

#CAlb_bg
#4hr overlapping DEGs (N=99): R2=0.7583, p<2.2e-16
#24hr overlapping DEGs (N=233): R2=0.04272, p<2.2e-16

#CAlb_man
#4hr overlapping DEGs (N=2): Not enough overlap
#24hr overlapping DEGs (N=41): R2=0.3351, p=4.391e-05
```

Overlap (Figure S15F)
```{r}
tt <- "4hr"
cond <- "CAlb_bg" 
cond <- "CAlb_live" 
cond <- "CAlb_man" 

venn::venn(list(subset(Masterfile_Bruno_2019)$Transcript, subset(Masterfile, condition == "CAlb" & time == tt)$Transcript), ilabels = "counts")

conds <- c("CAlb_bg", "CAlb_live", "CAlb_man")
times <- c("4hr", "24hr")

for (tt in times) {
  for (cond in conds) {
    bruno <- subset(Masterfile_Bruno_2019, time == tt & condition == cond)
    quant <- subset(Masterfile, condition == "CAlb" & time == tt)
    
    background <- intersect(subset(Masterfile_Bruno_2019, time == tt & condition == cond)$Transcript, subset(Masterfile, condition == "CAlb")$Transcript)
    de_both <- Reduce(intersect, list(subset(bruno, padj < 0.01)$Transcript, subset(quant, padj < 0.01)$Transcript))
    de_none <- background[!background %in% c(subset(bruno, padj < 0.01)$Transcript, subset(quant, padj < 0.01)$Transcript)]
    de_only_quant <- subset(quant, padj < 0.01)$Transcript[!subset(quant, padj < 0.01)$Transcript %in% subset(bruno, padj < 0.01)$Transcript]
    de_only_bruno <- subset(bruno, padj < 0.01)$Transcript[!subset(bruno, padj < 0.01)$Transcript %in% subset(quant, padj < 0.01)$Transcript]
    test <- data.frame(
      c(length(de_both), length(de_only_bruno)),
      c(length(de_only_quant), length(de_none))
    ) %>% fisher.test()
  }
}
```

## Bruno et al. 2020 (Figure S15G-H)
PCA (Figure S15G)
```{r}
#The bruno 2020 data is available at PMID: 33510868
bruno_2020_raw <- read.delim(here::here("validations/data/bruno_2020_raw.csv"), sep = ";")
bruno_2020_metadat <- bruno_2020_raw[1:3, ]

#Generating a metadata file
meta <- matrix(ncol = 4, nrow = 63)
rownames(meta) <- colnames(bruno_2020_metadat)[2:ncol(bruno_2020_metadat)]
meta[, 1] <- colnames(bruno_2020_metadat)[2:ncol(bruno_2020_metadat)]
meta[, 2] <- as.matrix(bruno_2020_metadat[2, 2:ncol(bruno_2020_metadat)])
meta[, 3] <- as.matrix(bruno_2020_metadat[1, 2:ncol(bruno_2020_metadat)])
meta[, 4] <- as.matrix(bruno_2020_metadat[3, 2:ncol(bruno_2020_metadat)])
meta <- as.data.frame(meta)
colnames(meta) <- c("id", "stimulus", "time", "donor")
bruno_2020_metadat <- meta

#Making a nice counts matrix
bruno_2020_raw <- bruno_2020_raw[4:nrow(bruno_2020_raw), ]
rownames(bruno_2020_raw) <- bruno_2020_raw$X
bruno_2020_raw <- bruno_2020_raw %>% dplyr::select(-X)
bruno_2020_raw <- as.matrix(bruno_2020_raw)
class(bruno_2020_raw) <- "numeric"
bruno_2020_raw <- as.data.frame(bruno_2020_raw)

bruno_2020_metadat_4hr <- subset(bruno_2020_metadat, time == "4h")
bruno_2020_metadat_24hr <- subset(bruno_2020_metadat, time == "24h")

raw_4hr <- colnames(bruno_2020_raw) %in% bruno_2020_metadat_4hr$id %>% which()
raw_24hr <- colnames(bruno_2020_raw) %in% bruno_2020_metadat_24hr$id %>% which()

bruno_2020_raw_4hr <- bruno_2020_raw[, raw_4hr] 
bruno_2020_raw_24hr <- bruno_2020_raw[, raw_24hr] 

dds_bruno_2020_4hr <- DESeqDataSetFromMatrix(
  countData = bruno_2020_raw_4hr, 
  colData = bruno_2020_metadat_4hr, 
  design = ~ donor + stimulus
  ) %>% 
  keepFromddsHTSeq %>% 
  mRNA_function

dds_bruno_2020_4hr$group <- relevel(dds_bruno_2020_4hr$group, ref = "RPMI")
dds_bruno_2020_4hr <- DESeq(dds_bruno_2020_4hr)

resultsNames(dds_bruno_2020_4hr) # lists the coefficients

res_4hr <- results(dds_bruno_2020_4hr)

opt.cond.long <- c("group_Candida.albicans_vs_RPMI", 
                   "group_Aspergillus.fumigatus_vs_RPMI", 
                   "group_Rhizopus.oryzae_vs_RPMI")

# or to shrink log fold changes association with condition:
res_CAlb_4hr <- lfcShrink(dds_bruno_2020_4hr, coef=opt.cond.long[1], type="apeglm")
res_AFum_4hr <- lfcShrink(dds_bruno_2020_4hr, coef=opt.cond.long[2], type="apeglm")
res_ROrz_4hr <- lfcShrink(dds_bruno_2020_4hr, coef=opt.cond.long[3], type="apeglm")

res_CAlb_4hr <- res_CAlb_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "CAlb") %>% tibble::rownames_to_column("Transcript")
res_AFum_4hr <- res_AFum_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "AFum") %>% tibble::rownames_to_column("Transcript")
res_ROrz_4hr <- res_ROrz_4hr %>% as.data.frame() %>% mutate(time = "4hr") %>% mutate(condition = "ROrz") %>% tibble::rownames_to_column("Transcript")
res_list_Bruno_2020_4hr <- list(res_CAlb_4hr, res_AFum_4hr, res_ROrz_4hr)

dds_bruno_2020_24hr <- DESeqDataSetFromMatrix(
  countData = bruno_2020_raw_24hr, 
  colData = bruno_2020_metadat_24hr, 
  design = ~ donor + stimulus
  ) %>% 
  keepFromddsHTSeq %>% 
  mRNA_function


dds_bruno_2020_24hr$group <- relevel(dds_bruno_2020_24hr$group, ref = "RPMI")
dds_bruno_2020_24hr <- DESeq(dds_bruno_2020_24hr)

resultsNames(dds_bruno_2020_24hr) # lists the coefficients

res_24hr <- results(dds_bruno_2020_24hr)

opt.cond.long <- c("group_Candida.albicans_vs_RPMI", 
                   "group_Aspergillus.fumigatus_vs_RPMI", 
                   "group_Rhizopus.oryzae_vs_RPMI")

# or to shrink log fold changes association with condition:
#resLFC24hrLPS <-         logFoldShrinkage(dds_24hr, opt.cond.long[1])
res_CAlb_24hr <- lfcShrink(dds_bruno_2020_24hr, coef=opt.cond.long[1], type="apeglm")
res_AFum_24hr <- lfcShrink(dds_bruno_2020_24hr, coef=opt.cond.long[2], type="apeglm")
res_ROrz_24hr <- lfcShrink(dds_bruno_2020_24hr, coef=opt.cond.long[3], type="apeglm")

res_CAlb_24hr <- res_CAlb_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "CAlb") %>% tibble::rownames_to_column("Transcript")
res_AFum_24hr <- res_AFum_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "AFum") %>% tibble::rownames_to_column("Transcript")
res_ROrz_24hr <- res_ROrz_24hr %>% as.data.frame() %>% mutate(time = "24hr") %>% mutate(condition = "ROrz") %>% tibble::rownames_to_column("Transcript")
res_list_Bruno_2020_24hr <- list(res_CAlb_24hr, res_AFum_24hr, res_ROrz_24hr)

Masterfile_Bruno_2020 <- rbind(
  ldply(res_list_Bruno_2020_4hr, data.frame), 
  ldply(res_list_Bruno_2020_24hr, data.frame)) %>% 
  addUpDown

for (i in 1:nrow(Masterfile_Bruno_2020)) {
  Masterfile_Bruno_2020$geneName[i] <- subset(ref_annot, Gene_id == Masterfile_Bruno_2020$Transcript[i])$Gene_name
}

Masterfile_Bruno_2020_padj <- Masterfile_Bruno_2020 %>% getDEGs
```

DE correlations
```{r}
t_1 <- "4hr"
t_1 <- "24hr"
times <- c("4hr", "24hr")
for (t_1 in times) {
    t <- full_join(
      Masterfile_Bruno_2020 %>% subset(condition == "CAlb") %>% subset(time == t_1),
      Masterfile %>% subset(condition == "CAlb") %>% subset(time == t_1),
      by = "Transcript"
    ) %>% subset(!is.na(log2FoldChange.x)) %>% subset(!is.na(log2FoldChange.y)) %>% subset(!is.na(padj.x)) %>% subset(!is.na(padj.y))

    
    s <- subset(t, padj.x < 0.01 & padj.y < 0.01)
    test <- lm(s$log2FoldChange.x ~ s$log2FoldChange.y) 
}

#4hr overlapping DEGs (N=325): R2=0.8215, p<2.2e-16
#24hr overlapping DEGs (N=1215): R2=0.7613, p<2.2e-16
```

Overlap (Figure S15H)
```{r}
times <- c("4hr", "24hr")
for (tt in times) {
  bruno_calb_4hr <- subset(Masterfile_Bruno_2020, time == tt & condition == "CAlb")
  quant <- subset(Masterfile, condition == "CAlb" & time == tt)
  
  background <- intersect(subset(Masterfile_Bruno_2020, time == tt & condition == "CAlb")$Transcript, subset(Masterfile, condition == "CAlb")$Transcript)
  de_both <- Reduce(intersect, list(subset(bruno_calb_4hr, padj < 0.01)$Transcript, subset(quant, padj < 0.01)$Transcript))
  de_none <- background[!background %in% c(subset(bruno_calb_4hr, padj < 0.01)$Transcript, subset(quant, padj < 0.01)$Transcript)]
  de_only_quant <- subset(quant, padj < 0.01)$Transcript[!subset(quant, padj < 0.01)$Transcript %in% subset(bruno_calb_4hr, padj < 0.01)$Transcript]
  de_only_bruno <- subset(bruno_calb_4hr, padj < 0.01)$Transcript[!subset(bruno_calb_4hr, padj < 0.01)$Transcript %in% subset(quant, padj < 0.01)$Transcript]
  test <- data.frame(
    c(length(de_both), length(de_only_bruno)),
    c(length(de_only_quant), length(de_none))
  ) %>% fisher.test()
}
```

# Functional classification (Figure 4A-B, S6A-C and S14A)
Numbers (Figure S14A)
```{r}
nudge = 30
conditions <- Masterfile_padj %>%
  group_by(`time-point`, condition, direction, funct.class) %>%
  summarize(n = n()) %>% 
  mutate(type = paste0(`time-point`, "_", condition, "_", direction)) %>% 
  group_by(type) %>%
  mutate(total = sum(n)) %>%
  mutate(rel = n / total) %>% 
  fixTimeOrder %>% 
  fixConditionOrder %>% 
  mutate(n = ifelse(direction == "down", -n, n)) %>%
  mutate(rel = ifelse(direction == "down", -rel, rel)) %>% 
  mutate(total_pos = ifelse(direction == "down", -total-nudge, total+nudge))

#Custom condition order
conditions$funct.class <- factor(conditions$funct.class, levels = c("uncharacterized", "other", "imm.ass", "PID"))

conditions %>% 
  ggplot()+
  geom_col(aes(x = `time-point`, y = n, fill = funct.class))+
  geom_text(aes(x = `time-point`, y = total_pos, label = total), size = 5) +
  scale_fill_viridis_d(limits = c("uncharacterized", "other", "imm.ass", "PID")) +
  #scale_fill_viridis_d(begin = 0, end = 1)+
  labs(x = "Time-point", y = "Differentially expressed genes (N)") +
  facet_grid(~condition)+
  geom_hline(yintercept = 0, linetype="dashed", color = "grey")+
  theme_bw()+
  theme(legend.position="top", text=element_text(size=26))
```

Relative (Figure 4A)
```{r}
rel_funct_class <- Masterfile_padj %>%
  mutate(funct.class = ifelse(uniprot_rev_score %in% c(1, 2), "uncharacterized", funct.class)) %>%
  mutate(funct.class = ifelse(is.na(funct.class), "other", funct.class)) %>% 
  group_by(`time-point`, condition, funct.class) %>%
  summarize(n = n()) %>%
  group_by(condition, `time-point`) %>% 
  mutate(total_cond = sum(n)) %>% 
  mutate(rel = n / total_cond) %>% 
  fixTimeOrder %>% 
  fixConditionOrder %>% 
  dplyr::select(-n, -total_cond) %>% 
  ungroup() %>% 
  tidyr::complete(`time-point`, condition, funct.class, fill = list(rel = 0))


  t <- rel_funct_class %>% mutate(perc = 100*rel)
  t$funct.class <- factor(t$funct.class, levels = c("other", "imm.ass", "PID", "uncharacterized"))
  cond <- unique(t$condition)
  
  t %>% 
    ggplot(aes(x = perc, y = `time-point`, fill = funct.class)) +
    geom_col(width = 0.5) +
    #scale_x_discrete(limits = c(" ", "4hr","24hr")) +
    scale_fill_brewer("Blues") +
    #coord_polar("y") +
    labs(x = "Percentage of DEGs", y = "Time point") +
    facet_grid(condition~.) +
    theme_bw() +
    theme(text = element_text(size = 22))
```

## FMUG (Figure S6A-C)
```{r}
#FMUG data available at PMID 38546716
fmug <- read.csv(here::here("data/main_table.csv"))
fmug_cols <- read.csv(here::here("data/main_table_columns.csv"))

Masterfile_fmug <- left_join(Masterfile %>% dplyr::rename(gene_ensembl = Transcript), fmug, by = "gene_ensembl") %>% 
  mutate(char = ifelse(funct.class == "uncharacterized", F, T))
```

Statistics
```{r}
deg <- subset(Masterfile, deg == "deg")$Transcript %>% unique()
non_deg <- subset(Masterfile, deg == "non-deg")$Transcript %>% unique()
non_deg <- non_deg[!non_deg %in% deg]

fmug_annot <- fmug %>% mutate(de_quant = ifelse(gene_ensembl %in% deg, T,
                                                ifelse(gene_ensembl %in% non_deg, F, NA)))

fmug_annot %>% dplyr::count(de_quant)

nrow(fmug_annot)
100* (length(intersect(c(deg, non_deg), unique(fmug_annot$gene_ensembl))) / length(c(deg, non_deg)))
100* (length(intersect(deg, unique(fmug_annot$gene_ensembl))) / length(deg))
venn::venn(list(c(deg, non_deg), unique(fmug_annot$gene_ensembl)), ilabel = "counts", snames = "quant, fmug")
venn::venn(list(deg, non_deg, unique(fmug_annot$gene_ensembl)), ilabel = "counts", snames = "deg, non-deg, fmug")

length(deg[!deg %in% fmug_annot$gene_ensembl]) / length(deg)
length(deg[!non_deg %in% fmug_annot$gene_ensembl]) / length(non_deg)

subset(Masterfile_fmug, !is.na(n_pubs)) %>% nrow()

#Testing for a difference in publications between characterized and uncharacterized DEGs
Masterfile_fmug %>% 
  subset(deg == "deg") %>% 
  ggplot(aes(x = char, y = n_pubs)) +
  geom_boxplot()

na.omit(subset(Masterfile_fmug, deg == "deg" & char == T)$n_pubs) %>% mean()
na.omit(subset(Masterfile_fmug, deg == "deg" & char == T)$n_pubs) %>% median()
na.omit(subset(Masterfile_fmug, deg == "deg" & char == F)$n_pubs) %>% mean()
na.omit(subset(Masterfile_fmug, deg == "deg" & char == F)$n_pubs) %>% median()
wilcox.test(na.omit(subset(Masterfile_fmug, deg == "deg" & char == T)$n_pubs), 
            na.omit(subset(Masterfile_fmug, deg == "deg" & char == F)$n_pubs))

#Testing for a difference in publications between DEGs and all other genes
fmug_annot %>% 
  ggplot(aes(x = de_quant, y = n_pubs)) +
  geom_boxplot()

na.omit(fmug_annot$n_pubs[fmug_annot$de_quant == T]) %>% mean()
na.omit(fmug_annot$n_pubs[fmug_annot$de_quant == T]) %>% median()

na.omit(fmug_annot$n_pubs[fmug_annot$de_quant == F]) %>% mean()
na.omit(fmug_annot$n_pubs[fmug_annot$de_quant == F]) %>% median()

wilcox.test(na.omit(fmug_annot$n_pubs[fmug_annot$de_quant == T]), na.omit(fmug_annot$n_pubs[fmug_annot$de_quant == F]))
```

Overlap (Figure S6A-B)
```{r}
fmug_overlap <- venn::venn(list(unique(Masterfile$Transcript), unique(Masterfile_padj$Transcript), fmug$gene_ensembl), 
           ilabels = "counts", snames = "All genes, DEGs, FMUG", box = F, ggplot = T, sncs = 1.2, ilcs = 1.5)

ggsave(here::here("figures/fmug_overlap.pdf"), device = "pdf", height = 6, width = 8, plot = fmug_overlap)

fmug_overlap_char_unchar <- venn::venn(list(
  unique(subset(Masterfile_padj, uniprot_unchar == "uncharacterized")$Transcript), 
  unique(subset(Masterfile_padj, uniprot_unchar == "characterized")$Transcript), 
  fmug$gene_ensembl), 
           ilabels = "counts", snames = "Uncharacterized, Characterized, FMUG", box = F, ggplot = T, sncs = 1.2, ilcs = 1.5)
```

Violin plot (Figure S6C)
```{r}
colnames(Masterfile_fmug)
fmug_diff <- Masterfile_fmug %>% subset(deg == "deg") %>% 
  dplyr::select(geneName, characterized = char, funct.class, n_pubs) %>% unique()

fmug_diff %>% 
  ggplot(aes(x = characterized, y = log10(n_pubs + 1))) +
  geom_violin(aes(fill = characterized)) +
  geom_boxplot(width=0.1, alpha=0.2) +
  stat_compare_means() +
  labs(x = "Characterized", y = "log10(N publicactions + 1)") +
  theme_classic() +
  theme(text = element_text(size = 16), legend.position = "none")
```

Overlap uncharacterized genes (Figure 4B)
```{r}
LPS_unchar_24hr <- subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2)) %>% getGenes() 
PolIC_unchar_24hr <- subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2)) %>% getGenes() 
SAur_unchar_24hr <- subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2)) %>% getGenes() 
CAlb_unchar_24hr <- subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2)) %>% getGenes() 

LPS_unchar_4hr_up <- subset(Masterfile_padj, condition == "LPS" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes() 
SAur_unchar_4hr_up <- subset(Masterfile_padj, condition == "SAur" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes()
PolIC_unchar_4hr_up <- subset(Masterfile_padj, condition == "PolIC" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes()
CAlb_unchar_4hr_up <- subset(Masterfile_padj, condition == "CAlb" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes()

LPS_unchar_24hr_up <- subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes() 
SAur_unchar_24hr_up <- subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes()
PolIC_unchar_24hr_up <- subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes()
CAlb_unchar_24hr_up <- subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "up") %>% getGenes()

LPS_unchar_4hr_down <- subset(Masterfile_padj, condition == "LPS" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes() 
SAur_unchar_4hr_down <- subset(Masterfile_padj, condition == "SAur" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes()
PolIC_unchar_4hr_down <- subset(Masterfile_padj, condition == "PolIC" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes()
CAlb_unchar_4hr_down <- subset(Masterfile_padj, condition == "CAlb" & `time-point` == "4hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes()

LPS_unchar_24hr_down <- subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes() 
SAur_unchar_24hr_down <- subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes()
PolIC_unchar_24hr_down <- subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes()
CAlb_unchar_24hr_down <- subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & uniprot_rev_score %in% c(1, 2) & direction == "down") %>% getGenes()

ggarrange(
  venn::venn(list(
    LPS_unchar_4hr_up,
    CAlb_unchar_4hr_up,
    PolIC_unchar_4hr_up,
    SAur_unchar_4hr_up
  ), ggplot = TRUE,
  ilcs = 1.5,
  sncs = 1.5,
  ilabel = "counts",
  snames = "LPS, C. albicans, Poly(I:C), S. aureus",
  box = F,
  zcolor = as.vector(viridis::viridis_pal()(5)[2:5])),
  
  venn::venn(list(
    LPS_unchar_4hr_down,
    CAlb_unchar_4hr_down,
    PolIC_unchar_4hr_down,
    SAur_unchar_4hr_down
  ), ggplot = TRUE,
  ilcs = 1.5,
  sncs = 1.5,
  ilabel = "counts",
  snames = "LPS, C. albicans, Poly(I:C), S. aureus",
  box = F,
  zcolor = as.vector(viridis::viridis_pal()(5)[2:5])),
  
  venn::venn(list(
    LPS_unchar_24hr_up,
    CAlb_unchar_24hr_up,
    PolIC_unchar_24hr_up,
    SAur_unchar_24hr_up
  ), ggplot = TRUE,
  ilcs = 1.5,
  sncs = 1.5,
  ilabel = "counts",
  snames = "LPS, C. albicans, Poly(I:C), S. aureus",
  box = F,
  zcolor = as.vector(viridis::viridis_pal()(5)[2:5])),
  
  venn::venn(list(
    LPS_unchar_24hr_down,
    CAlb_unchar_24hr_down,
    PolIC_unchar_24hr_down,
    SAur_unchar_24hr_down
  ), ggplot = TRUE,
  ilcs = 1.5,
  sncs = 1.5,
  ilabel = "counts",
  snames = "LPS, C. albicans, Poly(I:C), S. aureus",
  box = F,
  zcolor = as.vector(viridis::viridis_pal()(5)[2:5]))
)
```

# Functional enrichment of DEGs (Figure 3D, S14G and S16A-B)
Enrichment analysis (Figure 3D)
```{r}
degs_4hr_core <- c( 
  Reduce(intersect, list(subset(Masterfile_padj, condition == "LPS" & `time-point` == "4hr" & direction == "up")$geneName,
                         subset(Masterfile_padj, condition == "CAlb" & `time-point` == "4hr" & direction == "up")$geneName,
                         subset(Masterfile_padj, condition == "SAur" & `time-point` == "4hr" & direction == "up")$geneName,
                         subset(Masterfile_padj, condition == "PolIC" & `time-point` == "4hr" & direction == "up")$geneName))
)
degs_24hr_core <- c( 
  Reduce(intersect, list(subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & direction == "up")$geneName,
                         subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & direction == "up")$geneName,
                         subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & direction == "up")$geneName,
                         subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & direction == "up")$geneName)),
  Reduce(intersect, list(subset(Masterfile_padj, condition == "LPS" & `time-point` == "24hr" & direction == "down")$geneName,
                         subset(Masterfile_padj, condition == "CAlb" & `time-point` == "24hr" & direction == "down")$geneName,
                         subset(Masterfile_padj, condition == "SAur" & `time-point` == "24hr" & direction == "down")$geneName,
                         subset(Masterfile_padj, condition == "PolIC" & `time-point` == "24hr" & direction == "down")$geneName))
)

immune_core_deg_4 <- fisherGeneList(core = degs_4hr_core, 
                                    other = unique(subset(Masterfile_padj, `time-point` == "4hr")$geneName), 
                                    geneset = imm.ass)
PID_core_deg_4 <- fisherGeneList(core = degs_4hr_core, 
                                 other = unique(subset(Masterfile_padj, `time-point` == "4hr")$geneName), 
                                 geneset = PID)
mouse_core_deg_4 <- fisherGeneList(core = degs_4hr_core, 
                                   other = unique(subset(Masterfile_padj, `time-point` == "4hr")$geneName), 
                                   geneset = mouse_immu_genes)

immune_core_deg_24 <- fisherGeneList(core = degs_24hr_core, 
                                     other = unique(subset(Masterfile_padj, `time-point` == "24hr")$geneName), 
                                     geneset = imm.ass)
PID_core_deg_24 <- fisherGeneList(core = degs_24hr_core, 
                                  other = unique(subset(Masterfile_padj, `time-point` == "24hr")$geneName), 
                                  geneset = PID)
mouse_core_deg_24 <- fisherGeneList(core = degs_24hr_core, 
                                    other = unique(subset(Masterfile_padj, `time-point` == "24hr")$geneName), 
                                    geneset = mouse_immu_genes)


core_deg_time <- list(immune_core_deg_4, 
                      PID_core_deg_4, 
                      mouse_core_deg_4,
                      immune_core_deg_24,
                      PID_core_deg_24,
                      mouse_core_deg_24)

core_deg_time_stats <- do.call(rbind,
        lapply(core_deg_time, 
               function(t) data.frame(p_val = t$p.value, conf.int_1 = t$conf.int[1], conf.int_2 = t$conf.int[2], estimate = t$estimate)))
core_deg_time_stats$set <- c("immune_core_deg_4", "PID_core_deg_4", "mouse_core_deg_4",
                             "immune_core_deg_24", "PID_core_deg_24", "mouse_core_deg_24")

core_deg_time_stats$time <- sapply(core_deg_time_stats$set, function(t) str_split(t, "_") %>% purrr::map_chr(.,4))
core_deg_time_stats$geneset <- sapply(core_deg_time_stats$set, function(t) str_split(t, "_") %>% purrr::map_chr(.,1))             


core_deg_time_stats$geneset[core_deg_time_stats$geneset == "PID"] <- "IEI"
core_deg_time_stats$geneset[core_deg_time_stats$geneset == "mouse"] <- "Mouse immune"
core_deg_time_stats$geneset[core_deg_time_stats$geneset == "immune"] <- "Immune"
core_deg_time_stats$geneset <- factor(core_deg_time_stats$geneset, levels = c("IEI", "Mouse immune", "Immune"))

core_deg_time_stats$time[core_deg_time_stats$time == "4"] <- "4 hours"
core_deg_time_stats$time[core_deg_time_stats$time == "24"] <- "24 hours"
core_deg_time_stats$time <- factor(core_deg_time_stats$time, levels = c("24 hours", "4 hours"))


n_tests <- nrow(core_deg_time_stats)
#Correcting the CI's to reflect the Bonferroni correction
alpha_adjusted <- 0.01 / n_tests
core_deg_time_stats$conf.int_1_adj <- core_deg_time_stats$conf.int_1 - (1 - alpha_adjusted) / 2
core_deg_time_stats$conf.int_2_adj <- core_deg_time_stats$conf.int_2 + (1 - alpha_adjusted) / 2

core_deg_time_stats %>% 
  mutate(padj = p_val * nrow(core_deg_time_stats)) %>% 
  mutate(padj = ifelse(padj > 1, 1, padj)) %>% 
  mutate(stars = stars.pval(padj)) %>% 
  ggplot(aes(x = estimate, y = geneset, color = time)) +
  geom_point(position = position_dodge(width = 0.9), size = 3.5) +
  geom_errorbarh(aes(xmax = conf.int_1_adj, xmin = conf.int_2_adj), size = .5, position = position_dodge(width = 0.9), height = 0) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  scale_color_manual(values = c("#72727e", "#b0bcb0")) +
  labs(x = "OR", y = "Geneset", color = "Time point") +
  theme_bw() +
  theme(text = element_text(size = 21), legend.position = "top")
```

HPO analysis (Figure 3E)
```{r}
core_4hr_up_gprof <- gprofiler2::gost(query = Reduce(intersect, list(subset(LPS, `time-point` == "4hr" & direction == "up")$geneName, subset(CAlb, `time-point` == "4hr" & direction == "up")$geneName, subset(PolIC, `time-point` == "4hr" & direction == "up")$geneName, subset(SAur, `time-point` == "4hr" & direction == "up")$geneName)), 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = "GO:BP", as_short_link = FALSE)
core_4hr_up_gprof$result
nrow(core_4hr_up_gprof$result)

core_24hr_up_gprof <- gprofiler2::gost(query = Reduce(intersect, list(subset(LPS, `time-point` == "24hr" & direction == "up")$geneName, subset(CAlb, `time-point` == "24hr" & direction == "up")$geneName, subset(PolIC, `time-point` == "24hr" & direction == "up")$geneName, subset(SAur, `time-point` == "24hr" & direction == "up")$geneName)), 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = "GO:BP", as_short_link = FALSE)
core_24hr_up_gprof$result
nrow(core_24hr_up_gprof$result)

core_24hr_down_gprof <- gprofiler2::gost(query = Reduce(intersect, list(subset(LPS, `time-point` == "24hr" & direction == "down")$geneName, subset(CAlb, `time-point` == "24hr" & direction == "down")$geneName, subset(PolIC, `time-point` == "24hr" & direction == "down")$geneName, subset(SAur, `time-point` == "24hr" & direction == "down")$geneName)), 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = "GO:BP", as_short_link = FALSE)
core_24hr_down_gprof$result
```

IEI enrichment (Figure S14G)
```{r}
quant_iuis <- readxl::read_xlsx(here::here("data/quantseq_IEI_classification.xlsx"), sheet = "Sheet1") #This list has been checked for all table 6 IUIS genes (11-02-2025)

table6_genes <- quant_iuis[grep("6", quant_iuis$table),]$gene_quant
nottable6_genes <- quant_iuis[!grepl("6", quant_iuis$table),]$gene_quant


LPS <- subset(Masterfile_padj, condition == "LPS")$geneName
CAlb <- subset(Masterfile_padj, condition == "CAlb")$geneName
PolIC <- subset(Masterfile_padj, condition == "PolIC")$geneName
SAur <- subset(Masterfile_padj, condition == "SAur")$geneName

all_DEG <- c(LPS, CAlb, PolIC, SAur) %>% unique()

table6_enrichment <- fisherGeneList(core = overlap, other = all_DEG, geneset = table6_genes, conf = 0.99)
nottable6_enrichment <- fisherGeneList(core = overlap, other = all_DEG, geneset = nottable6_genes, conf = 0.99)


table6_tests <- do.call(rbind, lapply(list(table6_enrichment, nottable6_enrichment), 
                                      function(t) data.frame(p_val = t$p.value, 
                                                             conf.int_1 = t$conf.int[1], 
                                                             conf.int_2 = t$conf.int[2], 
                                                             estimate = t$estimate))) %>% tibble::rownames_to_column("set") %>% 
  mutate(padj = p_val*2)

table6_tests$geneset <- c("IEI innate immunity", "IEI non-innate immunity")
table6_tests$geneset <- factor(table6_tests$geneset, levels = c("IEI innate immunity", "IEI non-innate immunity"))  

table6_tests_plot <- table6_tests %>% 
  mutate(label = paste0("Padj=", scales::scientific(padj, digits = 3))) %>% 
  mutate(lab_x = 0.05) %>% 
  ggplot(aes(y = geneset, fill = geneset, label = label)) +
  geom_col(aes(x = -log10(padj))) +
  geom_text(aes(x = lab_x), size = 2, hjust = 0) +
  geom_vline(xintercept = -log10(0.01)) +
  xlab(expression('-log'[10]*"Padj")) +
  ylab("Geneset") +
  xlim(0, 3) +
  theme_classic() +
  theme(legend.position = "none")
```

## Cell type specific expression  (Figure S16A-B)
```{r}
#Cell type specific genes can be obtained from dice https://dice-database.org/#
unique_genes_files <- list.files(here::here("data/unique_expression"))
unique_genes <- list()

for (file in unique_genes_files) {
  #file <- unique_genes_files[1]
  unique_genes <- rlist::list.append(
    unique_genes,
    read.csv(paste0(here::here("ext_data/DICE/unique_expression/"), file)) %>% mutate(file = file)
  )
}
unique_genes_full <- unique_genes
unique_genes <- lapply(unique_genes, function(x) return(x$Gene.name))
names(unique_genes) <- sapply(unique_genes_files, function(t) stringr::str_split(t, "[.]") %>% purrr::map_chr(.,1))

unique_genes_full <- do.call(rbind, unique_genes_full)
```

Enrichment (Figure S16A)
```{r}
dice_celltypes_gsea <- rbind(
  GSEASimple(dds_4hr, test = "LPS", control = "RPMI", pathways = unique_genes, time = "4hr"),
  GSEASimple(dds_4hr, test = "C_Albicans", control = "RPMI", pathways = unique_genes, time = "4hr"),
  GSEASimple(dds_4hr, test = "S_Aureus", control = "RPMI", pathways = unique_genes, time = "4hr"),
  GSEASimple(dds_4hr, test = "PolyIC", control = "RPMI", pathways = unique_genes, time = "4hr"),
  
  GSEASimple(dds_24hr, test = "LPS", control = "RPMI", pathways = unique_genes, time = "24hr"),
  GSEASimple(dds_24hr, test = "C_Albicans", control = "RPMI", pathways = unique_genes, time = "24hr"),
  GSEASimple(dds_24hr, test = "S_Aureus", control = "RPMI", pathways = unique_genes, time = "24hr"),
  GSEASimple(dds_24hr, test = "PolyIC", control = "RPMI", pathways = unique_genes, time = "24hr")
)

dice_celltypes_gsea <- dice_celltypes_gsea %>% 
    mutate(condition = ifelse(condition == "C_Albicans", glue("<i>C. albicans<i>"), 
                       ifelse(condition == "S_Aureus", glue("<i>S. aureus<i>"), 
                              ifelse(condition == "PolyIC", "Poly(I:C)",
                                     ifelse(condition == "LPS", "LPS", condition)))))

dice_celltypes_gsea$condition <- factor(dice_celltypes_gsea$condition, levels = c("LPS", "<i>C. albicans<i>", "Poly(I:C)", "<i>S. aureus<i>"))

dice_celltypes_gsea_plot <- dice_celltypes_gsea %>% 
  mutate(star_pval = gtools::stars.pval(padj)) %>% 
  mutate(star_pval = ifelse(padj > 0.05, "", star_pval)) %>% 
  mutate(time_point = factor(time_point, levels = c("4hr", "24hr"))) %>% 
  ggplot(aes(x = condition, y = pathway, fill = NES, label = star_pval)) +
  geom_tile() +
  geom_text() +
  labs(x = "Condition", y = "Gene set") +
  scale_fill_gradient2(high = muted("red"), mid = "white", low = muted("blue")) +
  facet_grid(~time_point) +
  theme_bw() +
  theme(text = element_text(size = 20), legend.position = "none", axis.text.x = ggtext::element_markdown(angle = 45, vjust = 1, hjust=1))
```

Functional enrichment of NK-specific genes (Figure S16B)
```{r}
dice_celltypes_gsea <- rbind(
  GSEASimple(dds_4hr, test = "LPS", control = "RPMI", pathways = unique_genes, time = "4hr"),
  GSEASimple(dds_4hr, test = "C_Albicans", control = "RPMI", pathways = unique_genes, time = "4hr"),
  GSEASimple(dds_4hr, test = "S_Aureus", control = "RPMI", pathways = unique_genes, time = "4hr"),
  GSEASimple(dds_4hr, test = "PolyIC", control = "RPMI", pathways = unique_genes, time = "4hr"),
  
  GSEASimple(dds_24hr, test = "LPS", control = "RPMI", pathways = unique_genes, time = "24hr"),
  GSEASimple(dds_24hr, test = "C_Albicans", control = "RPMI", pathways = unique_genes, time = "24hr"),
  GSEASimple(dds_24hr, test = "S_Aureus", control = "RPMI", pathways = unique_genes, time = "24hr"),
  GSEASimple(dds_24hr, test = "PolyIC", control = "RPMI", pathways = unique_genes, time = "24hr")
)

nk_leading_edge_all <- do.call(c, subset(dice_celltypes_gsea, time_point == "24hr" & pathway == "NK")$leadingEdge) %>% unique()
nk_leading_edge_all_gprof <- gprofiler2::gost(nk_leading_edge_all)

nk_leading_edge_all_gprof_plot <- nk_leading_edge_all_gprof$result %>% subset(term_size < 500 & term_size > 100) %>% subset(source == "GO:BP") %>% 
  mutate(log10p = -log10(p_value)) %>% 
  ggplot(aes(x = log10p, y = reorder(term_name, log10p), fill = source)) +
  geom_col() +
  xlab(expression('-log'[10]*"P")) +
  labs(y = "Term name", fill = "Source") +
  theme_classic() +
  theme(text = element_text(size = 16), axis.ticks.y = element_blank(), legend.position = "top")
```

# GWAS hits associated with pathogen response DEGs (Figure S7A-D)
Gathering the genes per trait reported in the GWAS catalog
```{r}
#The gwas hits are available from https://www.ebi.ac.uk/gwas/
associations <- read.table(here::here("data/associations_v1.0.2.txt"), quote="", fill=FALSE, sep = "\t", header = T)

genes_per_trait <- list()
traits <- unique(associations$MAPPED_TRAIT)

for (trait in traits) {
  t <- subset(associations, MAPPED_TRAIT == trait)
  
  snp_genes <- sapply(t$SNP_GENE_IDS, function(t) str_split(t, ", ")) %>% unlist()
  snp_genes <- snp_genes[snp_genes != ""]
  names(snp_genes) <- NULL
  genes <- snp_genes %>% unique()

  genes_per_trait <- rlist::list.append(genes_per_trait, genes)
  names(genes_per_trait)[length(genes_per_trait)] <- trait
}
genes_per_trait_stats <- lapply(genes_per_trait, length) %>% as.data.frame() %>% t() %>% as.data.frame() %>% tibble::rownames_to_column("id") %>% rename(n = V1)
mean(genes_per_trait_stats$n)
median(genes_per_trait_stats$n)
```

Number of genes per trait (Figure S7A)
```{r}
genes_per_trait_stats <- data.frame(genes = sapply(genes_per_trait, length)) 

min(genes_per_trait_stats$genes)
max(genes_per_trait_stats$genes)
mean(genes_per_trait_stats$genes)
median(genes_per_trait_stats$genes)

genes_per_trait_stats_n <- genes_per_trait_stats %>% 
  mutate(cut = cut(genes, breaks = c(-1, 0, 1, 10, 100, 1000, Inf))) %>% 
  count(cut)

genes_per_trait_stats_n$cut <- as.character(genes_per_trait_stats_n$cut)
genes_per_trait_stats_n$cut[genes_per_trait_stats_n$cut == "(-1,0]"] <- 0
genes_per_trait_stats_n$cut[genes_per_trait_stats_n$cut == "(0,1]"] <- 1
genes_per_trait_stats_n$cut <- factor(genes_per_trait_stats_n$cut, levels = c("0", "1", "(1,10]", "(10,100]", "(100,1e+03]", "(1e+03,Inf]"))

genes_per_trait_stats_binned <- genes_per_trait_stats_n %>% 
  ggplot(aes(x = cut, y = n, label = n)) + 
  geom_col() +
  geom_text(nudge_y = 100) +
  geom_vline(xintercept = 10) +
  labs(x = "Genes (N)", y = "Traits (N)") + 
  theme_classic() + theme(text = element_text(size = 16), axis.text.x = element_text(angle = 45, hjust = 1))
```

Enrichment tests
```{r}
#Doing the enrichment tests for each time point and condition separately
cond <- unique(Masterfile$condition)
time <- unique(Masterfile$`time-point`)
gwas_test_out <- list()

for (c in cond) {
  for (t in time) {
    # c <- cond[1]
    # t <- time[1]
    dat <-  subset(Masterfile, `time-point` == t & condition == c)
    deg <- unique(subset(dat, deg == "deg")$Transcript)
    non_deg <- unique(subset(dat, deg == "non-deg")$Transcript)
    non_deg <- non_deg[non_deg %ni% deg]
    
    test_out <- list()
    for (trait in names(genes_per_trait)) {
      #trait <- "body height"
      #trait <- names(genes_per_trait)[1]
      if (length(genes_per_trait[[trait]]) != 0) {
        test <- data.frame(
          c(length(deg[deg %in% genes_per_trait[[trait]]]), length(non_deg[non_deg %in% genes_per_trait[[trait]]])),
          c(length(deg[deg %ni% genes_per_trait[[trait]]]), length(non_deg[non_deg %ni% genes_per_trait[[trait]]]))
        ) %>% fisher.test()
        
        test_out <- rlist::list.append(test_out, data.frame(name = trait, p = test$p.value, OR = test$estimate, OR_1 = test$conf.int[1], OR_2 = test$conf.int[2]))
      }
    }
    
    test_out <- do.call(rbind, test_out) 
    test_out <- test_out %>% mutate(padj = p*nrow(test_out))
    
    gwas_test_out <- rlist::list.append(gwas_test_out, test_out)
    names(gwas_test_out)[length(gwas_test_out)] <- paste0(c, "_", t)
  }
}
gwas_test_out <- bind_rows(gwas_test_out, .id = "id")
```

Number of hits (Figure S7B)
```{r}
n_gwas <- gwas_test_out %>% 
  subset(padj_add < 0.01) %>% 
  group_by(condition, time) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  tidyr::complete(condition, time, fill = list(n = 0))

mean(subset(n_gwas, time == "4hr")$n)
mean(subset(n_gwas, time == "24hr")$n)

n_gwas %>% 
  mutate(time = as.factor(time)) %>% 
  mutate(time = fct_relevel(time, c("4hr", "24hr"))) %>%
  ggplot(aes(x = condition, label = n, group = time)) +
  geom_col(aes(y = n, fill = time), position = "dodge") +
  #facet_grid(~time) +
  labs(x = "Condition", y = "Hits (N)", fill = "Time point") +
  geom_text(aes(y = n+2), size = 5, position = position_dodge(width = .9)) +
  theme_classic() +
  theme(text = element_text(size = 16), legend.position = "top")
```

Overlap of hits (Figure S7C)
```{r}
gwas_test_out_signif <- subset(gwas_test_out, padj_add < 0.01)
venn::venn(lapply(split(gwas_test_out_signif, gwas_test_out_signif$id), function(t) return(t$name)))

overlap <- UpSetR::upset(UpSetR::fromList(lapply(split(gwas_test_out_signif, gwas_test_out_signif$id), function(t) return(t$name))),
              order.by = "freq", text.scale = 2, nsets = length(unique(gwas_test_out_signif$id)))
```

Enrichment plot (Figure S7D)
```{r}
gwas_test_out %>% subset(padj_add < 0.01) %>% subset(name %in% overlap_traits_all) %>% 
  mutate(time = factor(time, levels = c("4hr", "24hr"))) %>% 
  #melt(id.vars = c("condition", "time", "name", "padj", "OR"), measure.vars = c("OR_1", "OR_2")) %>% 
  ggplot(aes(x = OR, y = reorder(name, OR), color = condition)) +
  geom_errorbarh(aes(xmax = OR_1, xmin = OR_2, color = condition), linewidth = .5, height = .2, position=position_dodge(width = 0.9)) +
  #facet_wrap(time~condition, scales = "free", nrow = 2) +
  facet_wrap(time~., ncol = 3) +
  geom_point(position=position_dodge(width=0.9)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  labs(x = "OR", y = "GWAS trait", fill = "Condition") +
  scale_color_viridis_d() +
  theme_classic() +
  theme(text = element_text(size = 16))
```

# eQLTs associated with pathogen response DEGs (Figure S8A-B)
Gathering eQTLs from DICE
```{r}
#The dice data is available from https://dice-database.org/#
dice_files <- list.files(here::here("data/dice"))
dice <- list()
for (file in dice_files) {
  vcf <- read.vcfR(paste0(here::here("ext_data/DICE/eqtl/"), file), verbose = FALSE )
  t <- vcf@fix %>% as.data.frame()
  t$gene <- sapply(t$INFO, function(x) str_split(x, ";") %>% purrr::map_chr(.,1) %>% str_split(., "=") %>% purrr::map_chr(.,2))
  t$symbol <- sapply(t$INFO, function(x) str_split(x, ";") %>% purrr::map_chr(.,2) %>% str_split(., "=") %>% purrr::map_chr(.,2))
  t$pvalue <- sapply(t$INFO, function(x) str_split(x, ";") %>% purrr::map_chr(.,3) %>% str_split(., "=") %>% purrr::map_chr(.,2))
  t$beta <- sapply(t$INFO, function(x) str_split(x, ";") %>% purrr::map_chr(.,4) %>% str_split(., "=") %>% purrr::map_chr(.,2))
  
  dice <- rlist::list.append(dice, t)
}
names(dice) <- sapply(dice_files, function(x) str_split(x, "[.]") %>% purrr::map_chr(.,1))
```

Overlap (Figure S8A)
```{r}
dice <- readRDS(here::here("ext_data/DICE/eqtl/dice_list.rda"))
dice_genesets <- lapply(dice, function(t) return(unique(t$symbol)))
overlap <- UpSetR::upset(UpSetR::fromList(dice_genesets), order.by = "freq", nsets = length(dice_genesets), nintersects = 20, text.scale = 3, mb.ratio = c(0.5, 0.5))
```

Enrichment test (Figure S8B)
```{r}
cond <- c("LPS", "CAlb", "SAur", "PolIC")
time <- c("4hr", "24hr")
dice_enrichment_sep <- list()

for (c in cond) {
  # c <- "PolIC"
  # tt <- "4hr"
  for (tt in time) {
    degs <- Masterfile$geneName[Masterfile$deg == "deg" & Masterfile$`time-point` == tt & Masterfile$condition == c] %>% unique()
    non_degs <- Masterfile$geneName[Masterfile$deg == "non-deg" & Masterfile$`time-point` == tt & Masterfile$condition == c] %>% unique()
    dice_enrichment <- list()
    for (i in 1:length(dice)) {
      t <- dice[[i]]
      test <- fisher.test(
        data.frame(
          DE = c(length(degs[degs %in% t$symbol]), length(degs[!degs %in% t$symbol])),
          non_DE = c(length(non_degs[non_degs %in% t$symbol]), length(non_degs[!non_degs %in% t$symbol]))
        )
      )
      dice_enrichment <- rlist::list.append(dice_enrichment, test)
    }
    names(dice_enrichment) <- names(dice)
    
    
    dice_enrichment_stats <- do.call(rbind,
                                     lapply(dice_enrichment, 
                                            function(t) data.frame(p_val = t$p.value, 
                                                                   conf.int_1 = t$conf.int[1], 
                                                                   conf.int_2 = t$conf.int[2], 
                                                                   estimate = t$estimate))) %>% tibble::rownames_to_column("set")
    
    n_tests <- nrow(dice_enrichment_stats)
    
    out <- dice_enrichment_stats %>% 
      mutate(padj = p_val * n_tests) %>% mutate(condition = c) %>% mutate(time_point = tt)
    
    dice_enrichment_sep <- rlist::list.append(dice_enrichment_sep, out)
  }
}
saveRDS(dice_enrichment_sep, here::here("output/dice_enrichment_sep.rda"))


dice_enrichment_sep_df <- do.call(rbind, dice_enrichment_sep)

dice_enrichment_sep_df %>% 
  mutate(padj = ifelse(padj > 1, 1, padj)) %>% 
  mutate(star_pval = stars.pval(padj)) %>% 
  mutate(star_pval = ifelse(star_pval == ".", "", star_pval)) %>% 
  mutate(time_point = factor(time_point, levels = c("4hr", "24hr"))) %>% 
  ggplot(aes(x = condition, y = set)) +
  geom_tile(aes(fill = estimate)) +
  geom_text(aes(label = star_pval), nudge_y = -0.1, size = 3) +
  labs(x = "Stimulus", y = "eQTL cell type", fill = "OR") +
  scale_fill_distiller(palette = "Blues", direction = 0) +
  facet_grid(~time_point)
```

# Gene network analysis of core DEGs (Figure 3F-G)
Performed in cytoscape, see: overlapping_degs.cys

# Co-expression analysis (Figure 4C-D, S9A-G, S17A-C)
4 hour data: prep
```{r}
t <- dds_4hr_t 
pheno <- read.delim(here("seq_data/Phenodata.txt")) %>% subset(time == "4hrs") 
name <- "both_wRPMI"

dds <- DESeq2::DESeqDataSetFromMatrix(
  countData = t, # Our prepped data frame with counts
  colData = pheno, # Data frame with annotation for our samples
  design = ~1 # Here we are not specifying a model
)

dds_norm_4hr <- rlog(dds)

normalized_counts_4hr <- assay(dds_norm_4hr) %>%
  t() # Transpose this data

gsg <- goodSamplesGenes(normalized_counts_4hr)
gsg$allOK == T
```

24 hour data: prep
```{r}
t <- dds_24hr_t 
pheno <- read.delim(here("seq_data/Phenodata.txt")) %>% subset(time == "24hrs") 
name <- "both_wRPMI"

dds <- DESeq2::DESeqDataSetFromMatrix(
  countData = t, # Our prepped data frame with counts
  colData = pheno, # Data frame with annotation for our samples
  design = ~1 # Here we are not specifying a model
)

dds_norm_24hr <- rlog(dds)

normalized_counts_24hr <- assay(dds_norm_24hr) %>%
  t() # Transpose this data

gsg <-goodSamplesGenes(normalized_counts_24hr)
gsg$allOK == T
```

Both time point: sample tree (Figure S9A)
```{r}
t <- dds_both 
pheno <- read.delim(here("seq_data/Phenodata.txt"))
name <- "both_wRPMI"

dds <- DESeq2::DESeqDataSetFromMatrix(
  countData = t, # Our prepped data frame with counts
  colData = pheno, # Data frame with annotation for our samples
  design = ~1 # Here we are not specifying a model
)

keep <- rowSums(counts(dds)>=1) >= 50
dds <- dds[keep,]

dds_norm_both <- rlog(dds)

normalized_counts_both <- assay(dds_norm_both) %>%
  t() # Transpose this data
saveRDS(normalized_counts_both, here::here("co-expression/output/normalized_counts_both.rda"))

gsg <-goodSamplesGenes(normalized_counts_both)
gsg$allOK == T

sampleTree <- hclust(dist(normalized_counts_both), method = "average")

plot(sampleTree, main = "", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
plot(sampleTree, main = "", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

4 hour data: sample tree (Figure S9B)
```{r}
normalized_counts_4hr <- readRDS(here::here("co-expression/output/normalized_counts_4hr.rda"))
sampleTree <- hclust(dist(normalized_counts_4hr), method = "average")
plot(sampleTree, main = "", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

24 hour data: sample tree (Figure S9C)
```{r}
normalized_counts_24hr <- readRDS(here::here("co-expression/output/normalized_counts_24hr.rda"))
sampleTree <- hclust(dist(normalized_counts_24hr), method = "average")
plot(sampleTree, main = "", sub="", xlab="", cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
```

Selecting a soft cutoff (Figure S9D-G)
```{r}
spt_4hr <- pickSoftThreshold(normalized_counts_4hr, dataIsExpr = TRUE, corFnc = cor, networkType = "signed")
spt_24hr <- pickSoftThreshold(normalized_counts_24hr, dataIsExpr = TRUE, corFnc = cor, networkType = "signed")

soft_threshold_4hr <- spt_4hr$fitIndices %>% 
  ggplot(aes(x = Power, y = `SFT.R.sq`, label = Power)) +
  geom_text() +
  labs(x="Soft Threshold (power)", y="Scale Free Topology Model Fit,signed R^2") +
  ggtitle("4hr soft threshold choice") +
  geom_hline(yintercept=0.8, color = "red", linewidth=1)

soft_threshold_24hr <- spt_24hr$fitIndices %>% 
  ggplot(aes(x = Power, y = `SFT.R.sq`, label = Power)) +
  geom_text() +
  labs(x="Soft Threshold (power)", y="Scale Free Topology Model Fit,signed R^2") +
  ggtitle("24hr soft threshold choice") +
  geom_hline(yintercept=0.8, color = "red", linewidth=1)

connectivity_4hr <- spt_4hr$fitIndices %>% 
  ggplot(aes(x = Power, y = `mean.k.`, label = Power)) +
  geom_text() +
  labs(x="Soft Threshold (power)", y="Mean Connectivity") +
  ggtitle("4hr connectivity") 

connectivity_24hr <- spt_24hr$fitIndices %>% 
  ggplot(aes(x = Power, y = `mean.k.`, label = Power)) +
  geom_text() +
  labs(x="Soft Threshold (power)", y="Mean Connectivity") +
  ggtitle("24hr connectivity") 

ggarrange(
  soft_threshold_4hr, soft_threshold_24hr,
  connectivity_4hr, connectivity_24hr
)
```

Generating the network modules: 4 hours
```{r}
normalized_counts_4hr <- readRDS(here::here("co-expression/output/normalized_counts_4hr.rda"))
bwnet_4hr <- blockwiseModules(normalized_counts_4hr,
  maxBlockSize = 5000, # What size chunks (how many genes) the calculations should be run in
  TOMType = "signed", # topological overlap matrix
  power = 5, # soft threshold for network construction
  numericLabels = FALSE,
  minModuleSize = 30,
  mergeCutHeight = 0.25,
  deepSplit = 2,
  randomSeed = 1234 # set a seed
)

gprof_4hr_modules <- list()
for (colour in unique(gene_module_key_4hr$module)) {
  go <- gprofiler2::gost(query = gene_module_key_4hr$gene[gene_module_key_4hr$module == colour], 
                         organism = "hsapiens", ordered_query = FALSE, 
                         multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                         measure_underrepresentation = FALSE, evcodes = FALSE, 
                         user_threshold = 0.05, correction_method = "g_SCS", 
                         domain_scope = "annotated", custom_bg = NULL, 
                         numeric_ns = "", sources = NULL, as_short_link = FALSE)
  if (!is.null(go)) {
      gprof_4hr_modules <- rlist::list.append(gprof_4hr_modules, go$result)
      if (is.null(names(gprof_4hr_modules))) {
        names(gprof_4hr_modules) <- colour
      } else {
        names(gprof_4hr_modules)[length(names(gprof_4hr_modules))] <- colour 
      }
  }
}
gprof_4hr_modules <- do.call(rbind, Map(cbind, gprof_4hr_modules, module=names(gprof_4hr_modules)))
```

Generating the network modules: 24 hours
```{r}
normalized_counts_24hr <- readRDS(here::here("output/normalized_counts_24hr.rda"))
bwnet_24hr <- blockwiseModules(normalized_counts_24hr,
  maxBlockSize = 5000, # What size chunks (how many genes) the calculations should be run in
  TOMType = "signed", # topological overlap matrix
  power = 10, # soft threshold for network construction
  numericLabels = FALSE,
  minModuleSize = 30,
  mergeCutHeight = 0.25,
  deepSplit = 2,
  randomSeed = 1234 # set a seed
)

gene_module_key_24hr <- tibble::enframe(bwnet_24hr$colors, name = "gene", value = "module") %>%
  # Let's add the `ME` part so its more clear what these numbers are and it matches elsewhere
  dplyr::mutate(module = paste0("ME", module))

gprof_24hr_modules <- list()
for (colour in unique(gene_module_key_24hr$module)) {
  go <- gprofiler2::gost(query = gene_module_key_24hr$gene[gene_module_key_24hr$module == colour], 
                         organism = "hsapiens", ordered_query = FALSE, 
                         multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                         measure_underrepresentation = FALSE, evcodes = FALSE, 
                         user_threshold = 0.05, correction_method = "g_SCS", 
                         domain_scope = "annotated", custom_bg = NULL, 
                         numeric_ns = "", sources = NULL, as_short_link = FALSE)
  if (!is.null(go)) {
      gprof_24hr_modules <- rlist::list.append(gprof_24hr_modules, go$result)
      if (is.null(names(gprof_24hr_modules))) {
        names(gprof_24hr_modules) <- colour
      } else {
        names(gprof_24hr_modules)[length(names(gprof_24hr_modules))] <- colour 
      }
  }
}
gprof_24hr_modules <- do.call(rbind, Map(cbind, gprof_24hr_modules, module=names(gprof_24hr_modules)))
```

GSEA using WGCNA modules (Figure 4C)
```{r}
WGCNAtoGSEA <- function(t, test, control, pathways, gsea_name) {
  df <- t %>% 
    results(contrast = c("group", test, control)) %>% 
    as.matrix()
  
  names <- rownames(df)
  rownames(df) <- NULL
  df <- cbind(names, df) %>% as.data.frame()
  df$stat <- as.numeric(df$stat)
  
  ranks <- df %>% dplyr::select(names, stat) %>% na.omit() %>% dplyr::distinct() %>% deframe() 
  
  fgseaRes <- fgsea::fgsea(pathways = pathways, 
                           stats = ranks, 
                           eps = 0)

  
  fgseaResTidy <- fgseaRes %>%
    as_tibble() %>%
    arrange(desc(NES))
  
  table <- fgseaResTidy %>% 
    dplyr::select(pathway, pval, padj, log2err, ES, NES, size) %>% 
    arrange(padj)
  return(table)
}

gene_module_4hr_list <- gene_module_key_4hr %>% group_by(module) %>% split(gene_module_key_4hr$module)
names(gene_module_4hr_list) <- sapply(names(gene_module_4hr_list), function(t) paste0(t, "_", "4hr"))
gene_module_24hr_list <- gene_module_key_24hr %>% group_by(module) %>% split(gene_module_key_24hr$module)
names(gene_module_24hr_list) <- sapply(names(gene_module_24hr_list), function(t) paste0(t, "_", "24hr"))

pathways = lapply(gene_module_4hr_list, function(t) return(t$gene))
  
wgcna_4hr_lps_4hr_modules <- WGCNAtoGSEA(dds_4hr, test = "LPS", control = "RPMI", gsea_name = "gsea_4hr", pathways = pathways)
wgcna_4hr_calb_4hr_modules <- WGCNAtoGSEA(dds_4hr, test = "C_Albicans", control = "RPMI", gsea_name = "gsea_4hr", pathways = pathways)
wgcna_4hr_saur_4hr_modules <- WGCNAtoGSEA(dds_4hr, test = "S_Aureus", control = "RPMI", gsea_name = "gsea_4hr", pathways = pathways)
wgcna_4hr_polic_4hr_modules <- WGCNAtoGSEA(dds_4hr, test = "PolyIC", control = "RPMI", gsea_name = "gsea_4hr", pathways = pathways)


pathways = lapply(gene_module_24hr_list, function(t) return(t$gene))

wgcna_24hr_lps_24hr_modules <- WGCNAtoGSEA(dds_24hr, test = "LPS", control = "RPMI", gsea_name = "gsea_24hr", pathways = pathways)
wgcna_24hr_calb_24hr_modules <- WGCNAtoGSEA(dds_24hr, test = "C_Albicans", control = "RPMI", gsea_name = "gsea_24hr", pathways = pathways)
wgcna_24hr_saur_24hr_modules <- WGCNAtoGSEA(dds_24hr, test = "S_Aureus", control = "RPMI", gsea_name = "gsea_24hr", pathways = pathways)
wgcna_24hr_polic_24hr_modules <- WGCNAtoGSEA(dds_24hr, test = "PolyIC", control = "RPMI", gsea_name = "gsea_24hr", pathways = pathways)

wgcna_gsea <- list(
  wgcna_4hr_lps_4hr_modules, wgcna_4hr_calb_4hr_modules, wgcna_4hr_saur_4hr_modules, wgcna_4hr_polic_4hr_modules,
  wgcna_24hr_lps_24hr_modules, wgcna_24hr_calb_24hr_modules, wgcna_24hr_saur_24hr_modules, wgcna_24hr_polic_24hr_modules
)
names(wgcna_gsea)
names(wgcna_gsea) <- c("wgcna_4hr_lps_4hr_modules", "wgcna_4hr_calb_4hr_modules", "wgcna_4hr_saur_4hr_modules", "wgcna_4hr_polic_4hr_modules", 
                       "wgcna_24hr_lps_24hr_modules", "wgcna_24hr_calb_24hr_modules", "wgcna_24hr_saur_24hr_modules", "wgcna_24hr_polic_24hr_modules")

map_df(wgcna_gsea, ~as.data.frame(.x), .id="id") %>% 
  mutate(time_dds = stringr::str_split(id, "_") %>% purrr::map_chr(.,2)) %>% 
  mutate(condition_dds = stringr::str_split(id, "_") %>% purrr::map_chr(.,3)) %>% 
  mutate(modules_time = stringr::str_split(id, "_") %>% purrr::map_chr(.,4)) %>% 
  subset(time_dds == modules_time) %>% 
  mutate(p_star = stars.pval(padj)) %>% 
  ggplot(aes(x = id, y = pathway, fill = NES, label = p_star)) +
  geom_tile() +
  geom_text() +
  facet_wrap(~modules_time, scales = "free") +
  scale_fill_distiller(type = "div", palette = 4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        text = element_text(size = 21))
```

Functional enrichment of gene modules (Figure 4D)
```{r}
go <- gprofiler2::gost(query = gene_module_key_4hr$gene[gene_module_key_4hr$module == "MEred"], 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = NULL, as_short_link = FALSE)
go$result %>% 
  subset(term_size >= 100 & term_size <= 500) %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC")) %>% 
  ggplot(aes(fill = source, x = -log10(p_value), y =reorder(term_name, -log10(p_value)))) +
  geom_col(width = 0.5) +
  theme_classic() +
  ggtitle("MEred_4hr")

go <- gprofiler2::gost(query = gene_module_key_4hr$gene[gene_module_key_4hr$module == "MEgreen"], 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = NULL, as_short_link = FALSE)
go$result %>% 
  subset(term_size >= 100 & term_size <= 500) %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC")) %>% 
  ggplot(aes(fill = source, x = -log10(p_value), y =reorder(term_name, -log10(p_value)))) +
  geom_col(width = 0.5) +
  theme_classic() +
  ggtitle("MEgreen_4hr")

go <- gprofiler2::gost(query = gene_module_key_24hr$gene[gene_module_key_24hr$module == "MEyellow"], 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = NULL, as_short_link = FALSE)
go$result %>% 
  subset(term_size >= 100 & term_size <= 500) %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC")) %>% 
  ggplot(aes(fill = source, x = -log10(p_value), y =reorder(term_name, -log10(p_value)))) +
  geom_col(width = 0.5) +
  theme_classic() +
  ggtitle("MEyellow_24hr")

go <- gprofiler2::gost(query = gene_module_key_24hr$gene[gene_module_key_24hr$module == "MEred"], 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = NULL, as_short_link = FALSE)
go$result %>% 
  subset(term_size >= 100 & term_size <= 500) %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC")) %>% 
  ggplot(aes(fill = source, x = -log10(p_value), y =reorder(term_name, -log10(p_value)))) +
  geom_col(width = 0.5) +
  theme_classic() +
  ggtitle("MEred_24hr")

go <- gprofiler2::gost(query = gene_module_key_24hr$gene[gene_module_key_24hr$module == "MEgreen"], 
                       organism = "hsapiens", ordered_query = FALSE, 
                       multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                       measure_underrepresentation = FALSE, evcodes = FALSE, 
                       user_threshold = 0.05, correction_method = "g_SCS", 
                       domain_scope = "annotated", custom_bg = NULL, 
                       numeric_ns = "", sources = NULL, as_short_link = FALSE)
go$result %>% 
  subset(term_size >= 100 & term_size <= 500) %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC")) %>% 
  ggplot(aes(fill = source, x = -log10(p_value), y =reorder(term_name, -log10(p_value)))) +
  geom_col(width = 0.5) +
  theme_classic() +
  ggtitle("MEgreen_24hr")
```

Number of genes per module (Figure S17A)
```{r}
rbind(
  gene_module_key_4hr %>% group_by(module) %>% summarize(n = n()) %>% mutate(time = "4hr"),
  gene_module_key_24hr %>% group_by(module) %>% summarize(n = n()) %>% mutate(time = "24hr")
) %>% 
  mutate(color = str_split(module, "ME") %>% map_chr(.,2)) %>%
  mutate(time = as.factor(time)) %>% 
  mutate(time = fct_relevel(time, c("4hr", "24hr"))) %>%
  ggplot(aes(x = module, y = n, label = n)) +
  facet_grid(~time) +
  labs(y = "Genes (N)", x = "WGCNA gene module") +
  geom_col(aes(fill = color)) +
  geom_text(nudge_y = 60, size = 5) +
  scale_fill_identity() +
  theme_bw() +
  theme(text = element_text(size = 16))
```

Functional enrichment of 4hr gene modules (Figure S17B)
```{r}
gprof_4hr_modules <- list()
for (colour in unique(gene_module_key_4hr$module)) {
  go <- gprofiler2::gost(query = gene_module_key_4hr$gene[gene_module_key_4hr$module == colour], 
                         organism = "hsapiens", ordered_query = FALSE, 
                         multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                         measure_underrepresentation = FALSE, evcodes = FALSE, 
                         user_threshold = 0.05, correction_method = "g_SCS", 
                         domain_scope = "annotated", custom_bg = NULL, 
                         numeric_ns = "", sources = NULL, as_short_link = FALSE)
  if (!is.null(go)) {
      gprof_4hr_modules <- rlist::list.append(gprof_4hr_modules, go$result)
      if (is.null(names(gprof_4hr_modules))) {
        names(gprof_4hr_modules) <- colour
      } else {
        names(gprof_4hr_modules)[length(names(gprof_4hr_modules))] <- colour 
      }
  }
}
gprof_4hr_modules <- do.call(rbind, Map(cbind, gprof_4hr_modules, module=names(gprof_4hr_modules)))

gprof_4hr_modules %>% subset(term_size > 100 & term_size < 500) %>% mutate(time = "4hr") %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC", "REAC", "CORUM", "HPA", "KEGG")) %>% 
  mutate(log10p = -log10(p_value)) %>% 
  ggplot(aes(y = tidytext::reorder_within(term_name, log10p, within = module), x = log10p, fill = source)) +
  geom_col() +
  ggtitle("WGCNA modules using 4hr stimulation data") +
  labs(x = "-log10(P)", y = "Gene set", fill = "Gene set source") +
  tidytext::scale_y_reordered() +
  ggh4x::facet_grid2(module~time, scales = "free_y", space = "free_y")
```

Functional enrichment of 4hr gene modules (Figure S17C)
```{r}
gprof_24hr_modules <- list()
for (colour in unique(gene_module_key_24hr$module)) {
  go <- gprofiler2::gost(query = gene_module_key_24hr$gene[gene_module_key_24hr$module == colour], 
                         organism = "hsapiens", ordered_query = FALSE, 
                         multi_query = FALSE, significant = T, exclude_iea = FALSE, 
                         measure_underrepresentation = FALSE, evcodes = FALSE, 
                         user_threshold = 0.05, correction_method = "g_SCS", 
                         domain_scope = "annotated", custom_bg = NULL, 
                         numeric_ns = "", sources = NULL, as_short_link = FALSE)
  if (!is.null(go)) {
      gprof_24hr_modules <- rlist::list.append(gprof_24hr_modules, go$result)
      if (is.null(names(gprof_24hr_modules))) {
        names(gprof_24hr_modules) <- colour
      } else {
        names(gprof_24hr_modules)[length(names(gprof_24hr_modules))] <- colour 
      }
  }
}
gprof_24hr_modules <- do.call(rbind, Map(cbind, gprof_24hr_modules, module=names(gprof_24hr_modules)))

gprof_24hr_modules %>% subset(term_size > 100 & term_size < 500) %>% mutate(time = "24hr") %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC", "REAC", "CORUM", "HPA", "KEGG")) %>% 
  mutate(log10p = -log10(p_value)) %>% 
  subset(source %in% c("GO:BP", "GO:MF", "GO:CC")) %>% 
  ggplot(aes(y = tidytext::reorder_within(term_name, log10p, within = module), x = log10p, fill = source)) +
  geom_col() +
  ggtitle("WGCNA modules using 24hr stimulation data") +
  labs(x = "-log10(P)", y = "Gene set", fill = "Gene set source") +
  tidytext::scale_y_reordered() +
  ggh4x::facet_grid2(module~time, scales = "free_y", space = "free_y")
```

Gene networks (Figure 4D)
Figure produced in cytoscape, see: wgcna_correlations.cys
```{r}
networks_filtered <- list.files(here::here("co-expression/output/networks/"), "filtered.tsv")
networks <- list.files(here::here("co-expression/output/networks/"), ".tsv")
networks_raw <- networks[networks %ni% networks_filtered]

hist(yellow_4hr$correlation)
subset(yellow_4hr, correlation > 0.05)$gene1 %>% unique() %>% length()

gene_module_key_4hr$module[gene_module_key_4hr$gene %in% unchar_degs_ens] %>% unique()

#I want to get the uncharacterized genes too
any(unchar_degs_ens %in% unique(subset(yellow_4hr, correlation > 0.02)$gene1))
subset(subset(yellow_4hr, correlation > 0.02), gene1 %in% unchar_degs_ens)$gene1 %>% unique() 
#There are two uncharacterized genes in there

gprofiler2::gconvert(unique(subset(yellow_4hr, correlation > 0.02)$gene1))$name %>% unique() %>% clipr::write_clip()

#These are the genenames of the unchar genes in this module
gprofiler2::gconvert(subset(yellow_4hr, gene1 %in% unchar_degs_ens)$gene1)$name %>% unique() 
gprofiler2::gconvert(subset(yellow_4hr, correlation > 0.02 & gene1 %in% unchar_degs_ens)$gene1)$name %>% unique() 

out <- subset(yellow_4hr, correlation > 0.02)
```

Example DEGs log2foldchange (Figure 4E)
```{r}
example <- Masterfile %>% mutate(star_p = stars.pval(padj)) %>% mutate(star_p = ifelse(padj > 0.05 | is.na(padj), "ns", star_p)) %>% mutate(`time-point` = factor(`time-point`, levels = c("4hr", "24hr"))) %>% 
  subset(geneName %in% c("KIAA0040", "FAM49A")) %>% 
  mutate(ci_1 = log2FoldChange-lfcSE, ci_2 = log2FoldChange+lfcSE) %>% 
    mutate(condition = ifelse(condition == "CAlb", glue("<i>C. albicans<i>"), 
                            ifelse(condition == "SAur", glue("<i>S. aureus<i>"), 
                                   ifelse(condition == "PolIC", "Poly(I:C)",
                                          ifelse(condition == "LPS", "LPS", condition))))) %>% 
  mutate(geneName = ifelse(geneName == "FAM49A", glue("<i>FAM49A<i>"),
                           ifelse(geneName == "KIAA0040", glue("<i>KIAA0040<i>"), geneName)))

example$condition <- factor(example$condition, 
                            levels = c("LPS", "<i>C. albicans<i>", "Poly(I:C)", "<i>S. aureus<i>"))

example %>% 
  ggplot(aes(y = `time-point`, label = paste0("LFC=", round(log2FoldChange, 2), "; p=", scientific(pvalue, digits = 2)))) +
  geom_point(aes(x = log2FoldChange, color = `time-point`), size = 2, position=position_dodge(width = 0.9), stat = "identity")  +
  labs(x = "Log2FoldChange", y = "Time-point", color = "Time-point") +
  geom_errorbarh(aes(xmax = ci_1, xmin = ci_2, color = `time-point`), 
                 linewidth = .5, height = .1, position=position_dodge(width = 0.9)) +
  geom_text(aes(x = 1),nudge_y = 0.25, nudge_x = 0, size = 1.7) +
  facet_grid(geneName~condition) +
  theme_bw() +
  theme(strip.text = ggtext::element_markdown())
```

# Cell-type level expression of uncharacterized genes (Figure S18A-D)
CELLxGENE data (Figure S18A)
```{r}
#The cellxgene data is available at https://cellxgene.cziscience.com/gene-expressions
cxg_example <- read.csv(here::here("ext_data/cellxgene/CELLxGENE_gene_expression_090924.csv"), skip = 9) 
rownames(cxg_example) <- NULL
colnames(cxg_example) <- c("Tissue", "Cell_type", "Cell_count", "Tissue_comp", "Gene", "Expression", "Scaled_expression", "N_cells_expression")

cxg_example <- cxg_example %>% mutate(Perc_express = N_cells_expression / Cell_count)

cxg_example %>% 
  subset(Tissue == "blood") %>% 
  subset(Cell_count > 10^5) %>% 
  subset(!Cell_type %in% c("cell")) %>% 
  subset(Gene %in% c("KIAA0040", "CYRIA")) %>% 
  ggplot(aes(x = Gene, y = tidytext::reorder_within(Cell_type, Scaled_expression, within = Gene), 
             color = Scaled_expression, size = Perc_express)) + 
  geom_point() +
  tidytext::scale_y_reordered() +
  labs(x = "Gene", y = "Cell type", fill = "Gene expression", size = "Expressed in cells (%)") +
  facet_wrap(~Gene, scales = "free") +
  scale_color_viridis_c(option = "inferno", begin = 1, end = 0, limits = c(0, 1)) + 
  theme_classic() +
  theme(text = element_text(size = 16))
```

HPA immune (Figure S18B)
```{r}
#The hpa dataset is available at https://www.proteinatlas.org/about/download
hpa_immu <- read_tsv(here::here("ext_data/human_protein_atlas/immune/rna_immune_cell.tsv"))
hpa_immu %>% subset(`Gene name` %in% c("KIAA0040", "CYRIA")) %>% 
  subset(`Immune cell` != "total PBMC") %>% 
  mutate(`Gene name` = factor(`Gene name`, levels = c("KIAA0040", "CYRIA"))) %>% 
  ggplot(aes(y = tidytext::reorder_within(`Immune cell`, -pTPM, within = `Gene name`), x = pTPM)) +
  geom_col() +
  labs(y = "Cell type") +
  tidytext::scale_y_reordered() +
  facet_wrap(~`Gene name`, scales = "free") +
  ggtitle("HPA immune") +
  theme_classic() +
  theme(text = element_text(size = 16))
```

Monaco dataset (Figure S18C)
```{r}
#The monaco dataset is available from PMID: 30726743
monaco <- read_tsv(here::here("ext_data/human_protein_atlas/immune/rna_immune_cell_monaco.tsv"))
monaco %>% subset(`Gene name` %in% c("KIAA0040", "CYRIA")) %>% 
  mutate(`Gene name` = factor(`Gene name`, levels = c("KIAA0040", "CYRIA"))) %>% 
  subset(`Immune cell` != "total PBMC") %>% 
  ggplot(aes(y = tidytext::reorder_within(`Immune cell`, -pTPM, within = `Gene name`), x = pTPM)) +
  geom_col() +
  labs(y = "Cell type") +
  tidytext::scale_y_reordered() +
  facet_wrap(~`Gene name`, scales = "free") +
  ggtitle("Monaco et al.") +
  theme_classic() +
  theme(text = element_text(size = 16))
```

Schmiedel dataset (Figure S18D)
```{r}
#The schmiedel dataset is available from PMID: 30449622
schmiedel <- read_tsv(here::here("ext_data/human_protein_atlas/immune/rna_immune_cell_schmiedel.tsv"))
schmiedel %>% subset(`Gene name` %in% c("KIAA0040", "CYRIA")) %>% 
  mutate(`Gene name` = factor(`Gene name`, levels = c("KIAA0040", "CYRIA"))) %>% 
  subset(`Immune cell` != "total PBMC") %>% 
  ggplot(aes(y = tidytext::reorder_within(`Immune cell`, -TPM, within = `Gene name`), x = TPM)) +
  geom_col() +
  labs(y = "Cell type") +
  tidytext::scale_y_reordered() +
  facet_wrap(~`Gene name`, scales = "free") +
  ggtitle("Schmiedel et al.") +
  theme_classic() +
  theme(text = element_text(size = 16))
```

# Co-expression from published data (Figure S19A-B)
```{r}
TERM2GENE <- correlationAnalyzeR::getTERM2GENE(GSEA_Type = c("GO:BP")) 
pathways <- split(TERM2GENE, TERM2GENE$gs_name)
pathways <- lapply(pathways, function(t) return(t$gene_symbol))


gene <- "KIAA0040"
gene <- "FAM49A"

output <- correlationAnalyzeR::analyzeSingleGenes(gene, topPlots = F)
corr <- output$correlations %>% tibble::deframe()
names(corr) <- rownames(output$`P values`)
fgseaRes <- fgsea::fgsea(pathways = pathways, 
                  stats = corr,
                  min = 100,
                  max = 500)

topPathwaysUp <- fgseaRes[ES > 0][head(order(pval), n=10), pathway]
topPathways <- c(topPathwaysUp)
fgsea::plotGseaTable(pathways[topPathways], corr, fgseaRes, 
              gseaParam=0.5)
```